---
title: "Summaries"
author: "David Ory"
output:
  html_document:
    theme: cosmo
    toc: yes
---
## Administration

#### Purpose
Combines legacy data (see `Extract Variables from Legacy Surveys` and `Build Legacy Database`) with standard data (see `Build Standard Database`) and then extracts CSV files for use in Tableau summaries. 

#### _ISSUES_

#### _TODO_
1.  Churn through Tableau
2.  Write out public data set

## Overhead

#### Libraries
```{r overhead}
library(knitr)
suppressMessages(library(dplyr))
library(stringr)
library(reshape2)
```

#### Knitr config
```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote file names
```{r file-names}
F_INPUT_LEGACY_RDATA   = "M:/Data/OnBoard/Data and Reports/_data Standardized/survey_legacy.RData"
F_INPUT_STANDARD_RDATA = "M:/Data/OnBoard/Data and Reports/_data Standardized/survey_standard.RData"

D_OUTPUT_TABLEAU = "M:/Data/OnBoard/Data and Reports/_data Standardized/tableau/"

```

## Combine Legacy and Standard data
```{r combine}
load(F_INPUT_LEGACY_RDATA)
load(F_INPUT_STANDARD_RDATA)

str(survey.standard, list.len = 500)
str(survey.legacy, list.len = 500)

data.ready <- rbind(survey.standard, survey.legacy)

# Replace NA with missing
data.ready.sum <- data.ready %>%
  mutate(race             = ifelse(is.na(race),             'Missing', race)) %>%
  mutate(hispanic         = ifelse(is.na(hispanic),         'Missing', hispanic)) %>%
  mutate(household_income = ifelse(is.na(household_income), 'Missing', household_income)) %>%
  mutate(access_mode      = ifelse(is.na(access_mode),      'Missing', access_mode)) %>%
  mutate(egress_mode      = ifelse(is.na(egress_mode),      'Missing', egress_mode)) %>%
  mutate(boardings        = ifelse(is.na(boardings),        'Missing', boardings)) %>%
  mutate(depart_hour      = ifelse(is.na(depart_hour),      'Missing', depart_hour)) %>%
  mutate(return_hour      = ifelse(is.na(return_hour),      'Missing', return_hour)) %>%
  mutate(tour_purp        = ifelse(is.na(tour_purp),        'Missing', tour_purp)) %>%
  mutate(auto_suff        = ifelse(is.na(auto_suff),        'Missing', auto_suff)) %>%
  mutate(work_status      = ifelse(is.na(work_status),      'Missing', work_status)) %>%
  mutate(student_status   = ifelse(is.na(student_status),   'Missing', student_status))
  
```

## Tableau Summaries

### Race by Ethnicity by Income by Operator by Weekpart
```{r race-ethnicity-income}
sum.race <- data.ready.sum %>%
  group_by(operator, race, hispanic, household_income, weekpart) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.race, paste(D_OUTPUT_TABLEAU, "RaceHispanicByOperator.csv", sep = ""), row.names = FALSE, quote = T)
```

###  Access by Egress by Transfers by Operator by Weekpart
```{r acess-egress-transfers}
sum.tofro <- data.ready.sum %>%
  group_by(operator, weekpart, access_mode, egress_mode, boardings) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.tofro, paste(D_OUTPUT_TABLEAU, "AccessEgressTransfers.csv", sep = ""), row.names = FALSE, quote = T)

```

### Time Leaving Home by Time Returning Home by Tour Purpose by Operator by Weekpart
```{r depart-arrive-purp}
sum.time <- data.ready.sum %>%
  group_by(operator, weekpart, depart_hour, return_hour, tour_purp) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.time, paste(D_OUTPUT_TABLEAU, "DepartReturnPurpose.csv", sep = ""), row.names = FALSE, quote = T)

```

## Automobile Sufficiency by Worker Status by Student Status by Boardings by Operator by Weekpart
```{r autosuff-status-transfers}
sum.suff <- data.ready.sum %>%
  group_by(operator, weekpart, auto_suff, work_status, student_status, boardings) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.suff, paste(D_OUTPUT_TABLEAU, "AutoWorkStudent.csv", sep = ""), row.names = FALSE, quote = T)

```
