---
title: "Summaries"
author: "David Ory"
output:
  html_document:
    theme: cosmo
    toc: yes
---
## Administration

#### Purpose
Combines legacy data (see `Extract Variables from Legacy Surveys` and `Build Legacy Database`) with standard data (see `Build Standard Database`) and then extracts `CSV` files for use in Tableau summaries.  Script also writes out combined data sets to disk as both `RData` and `CSV`.

## Overhead

#### Libraries
```{r overhead}
library(knitr)
suppressMessages(library(dplyr))
library(stringr)
library(reshape2)
```

#### Knitr config
```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote file names
```{r file-names}
F_INPUT_LEGACY_RDATA   = "M:/Data/OnBoard/Data and Reports/_data Standardized/survey_legacy.RData"
F_INPUT_STANDARD_RDATA = "M:/Data/OnBoard/Data and Reports/_data Standardized/survey_standard.RData"

F_INPUT_COUNTY  = "county_to_taz_reference.csv"

F_OUTPUT_CSV   = "M:/Data/OnBoard/Data and Reports/_data Standardized/share_data/survey.csv"
F_OUTPUT_RDATA = "M:/Data/OnBoard/Data and Reports/_data Standardized/share_data/survey.RData"

D_OUTPUT_TABLEAU = "M:/Data/OnBoard/Data and Reports/_data Standardized/tableau/"
```

## Combine Legacy and Standard data
```{r combine}
load(F_INPUT_LEGACY_RDATA)
load(F_INPUT_STANDARD_RDATA)

str(survey.standard, list.len = 500)
str(survey.legacy,   list.len = 500)

# DEBUG
legacy_names <- colnames(survey.legacy)
std_names <- colnames(survey.standard)

for (std_name in std_names) {
  if(!(std_name %in% legacy_names))
    print(std_name)
}

for (legacy_name in legacy_names) {
  if(!(legacy_name %in% std_names))
    print(legacy_name)
}
# end DEBUG

data.ready <- rbind(survey.standard, survey.legacy)

# Remove the BART pre-test data
data.ready <- data.ready %>%
  filter(!(operator == "BART PRE-TEST"))

# Add location counties from TAZ geo-coding
county_ref <- read.csv(F_INPUT_COUNTY, header = TRUE, stringsAsFactors = FALSE)
home_join <- select(county_ref, home_taz = TAZ,      home_county = COUNTY)
work_join <- select(county_ref, workplace_taz = TAZ, workplace_county = COUNTY)
schl_join <- select(county_ref, school_taz = TAZ,    school_county = COUNTY)

data.ready <- left_join(data.ready, home_join, by = c("home_taz"))
data.ready <- left_join(data.ready, work_join, by = c("workplace_taz"))
data.ready <- left_join(data.ready, schl_join, by = c("school_taz"))

# Replace NA with missing
data.ready.sum <- data.ready %>%
  mutate(race             = ifelse(is.na(race),             'Missing', race)) %>%
  mutate(hispanic         = ifelse(is.na(hispanic),         'Missing', hispanic)) %>%
  mutate(household_income = ifelse(is.na(household_income), 'Missing', household_income)) %>%
  mutate(access_mode      = ifelse(is.na(access_mode),      'Missing', access_mode)) %>%
  mutate(egress_mode      = ifelse(is.na(egress_mode),      'Missing', egress_mode)) %>%
  mutate(boardings        = ifelse(is.na(boardings),        'Missing', boardings)) %>%
  mutate(depart_hour      = ifelse(is.na(depart_hour),      'Missing', depart_hour)) %>%
  mutate(return_hour      = ifelse(is.na(return_hour),      'Missing', return_hour)) %>%
  mutate(tour_purp        = ifelse(is.na(tour_purp),        'Missing', tour_purp)) %>%
  mutate(auto_suff        = ifelse(is.na(auto_suff),        'Missing', auto_suff)) %>%
  mutate(work_status      = ifelse(is.na(work_status),      'Missing', work_status)) %>%
  mutate(student_status   = ifelse(is.na(student_status),   'Missing', student_status)) %>%
  mutate(fare_medium      = ifelse(is.na(fare_medium),      'Missing', fare_medium)) %>%
  mutate(approximate_age  = ifelse(is.na(approximate_age),  'Missing', approximate_age)) %>%
  mutate(fare_category    = ifelse(is.na(fare_category),    'Missing', fare_category)) %>%
  mutate(eng_proficient   = ifelse(is.na(eng_proficient),   'Missing', eng_proficient))
  
```

## Data to Disk as CSV and RData
```{r write-disk}
# write.csv(data.ready, file = F_OUTPUT_CSV, row.names = FALSE, quote = T)
# 
# survey <- data.ready
# save(survey, file = F_OUTPUT_RDATA)
```


## Tableau Summaries

### Race by Ethnicity by Income by Operator by Weekpart
```{r race-ethnicity-income}
sum.race <- data.ready.sum %>%
  group_by(operator, race, hispanic, household_income, weekpart) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.race, paste(D_OUTPUT_TABLEAU, "RaceHispanicByOperator.csv", sep = ""), row.names = FALSE, quote = T)
```

###  Access by Egress by Transfers by Operator by Weekpart
```{r acess-egress-transfers}
sum.tofro <- data.ready.sum %>%
  group_by(operator, weekpart, access_mode, egress_mode, boardings) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.tofro, paste(D_OUTPUT_TABLEAU, "AccessEgressTransfers.csv", sep = ""), row.names = FALSE, quote = T)

```

### Time Leaving Home by Time Returning Home by Tour Purpose by Operator by Weekpart
```{r depart-arrive-purp}
sum.time <- data.ready.sum %>%
  group_by(operator, weekpart, depart_hour, return_hour, tour_purp) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.time, paste(D_OUTPUT_TABLEAU, "DepartReturnPurpose.csv", sep = ""), row.names = FALSE, quote = T)

```

### Automobile Sufficiency by Worker Status by Student Status by Boardings/Transfers by Operator by Weekpart
```{r autosuff-status-transfers}
sum.suff <- data.ready.sum %>%
  group_by(operator, weekpart, auto_suff, work_status, student_status, boardings) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.suff, paste(D_OUTPUT_TABLEAU, "AutoWorkStudent.csv", sep = ""), row.names = FALSE, quote = T)

```

### Age by Fare Media by Fare Category by Operator by Weekpart
```{r age-media-category}
sum.fare <- data.ready.sum %>%
  group_by(operator, weekpart, approximate_age, fare_medium, fare_category) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.fare, paste(D_OUTPUT_TABLEAU, "AgeMediumCategory.csv", sep = ""), row.names = FALSE, quote = T)

```

### Home County by Work County by School County by Work Status by Student Status by Operator by Weekpart
```{r home-work-school}
sum.locs <- data.ready.sum %>%
  group_by(operator, weekpart, home_county, workplace_county, school_county, work_status, student_status) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.locs, paste(D_OUTPUT_TABLEAU, "HomeWorkSchoolCounty.csv", sep = ""), row.names = FALSE, quote = T)

```

### Language at Home, Survey Language, English Proficiency by Operator and Weekpart
```{r language}
sum.lang <- data.ready.sum %>%
  group_by(operator, weekpart, interview_language, language_at_home, field_language, eng_proficient) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.lang, paste(D_OUTPUT_TABLEAU, "Language.csv", sep = ""), row.names = FALSE, quote = T)

```

### Field Dates and Day of Week
```{r field-dates-day-of-week}
sum.date <- data.ready.sum %>%
  group_by(operator, weekpart, field_start, field_end, day_of_the_week) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.date, paste(D_OUTPUT_TABLEAU, "FieldDates.csv", sep = ""), row.names = FALSE, quote = T)

```

### First Boarding and Last Alighting
```{r boarding-alighting}
sum.board <- data.ready.sum %>%
  filter(!is.na(first_board_lat)) %>%
  filter(!is.na(first_board_lon)) %>%
  filter(!is.na(last_alight_lat)) %>%
  filter(!is.na(last_alight_lon)) %>%
  group_by(operator, weekpart, day_part, first_board_lat, first_board_lon, last_alight_lat, last_alight_lon) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.board, paste(D_OUTPUT_TABLEAU, "BoardingAndAlighting.csv", sep = ""), row.names = FALSE, quote = T)

```

### Distribution of Weights
```{r weights}
sum.wgt <- data.ready.sum %>%
  select(ID, operator, weekpart, weight)

write.csv(sum.wgt, paste(D_OUTPUT_TABLEAU, "DistributionOfWeights.csv", sep = ""), row.names = FALSE, quote = T)

```
### Simplified Auto Sufficiency
```{r autos}
sum.autos <- data.ready.sum %>%
  group_by(operator, weekpart, auto_suff) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.autos, paste(D_OUTPUT_TABLEAU, "AutoSufficiency.csv", sep = ""), row.names = FALSE, quote = T)

```
### Gender
```{r gender}
sum.gender <- data.ready.sum %>%
  group_by(operator, weekpart, gender) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.gender, paste(D_OUTPUT_TABLEAU, "Gender.csv", sep = ""), row.names = FALSE, quote = T)

```

### Age Category
```{r agecat}

data.ready.sum$agecat[data.ready.sum$approximate_age>=0 & data.ready.sum$approximate_age <= 17]     <- "17 or Younger"
data.ready.sum$agecat[data.ready.sum$approximate_age >=18 & data.ready.sum$approximate_age <= 24]   <- "18 to 24"
data.ready.sum$agecat[data.ready.sum$approximate_age >=25 & data.ready.sum$approximate_age <= 34]   <- "25 to 34"
data.ready.sum$agecat[data.ready.sum$approximate_age >=35 & data.ready.sum$approximate_age <= 44]   <- "35 to 44"
data.ready.sum$agecat[data.ready.sum$approximate_age >=45 & data.ready.sum$approximate_age <= 54]   <- "45 to 54"
data.ready.sum$agecat[data.ready.sum$approximate_age >=55 & data.ready.sum$approximate_age <= 64]   <- "55 to 64"
data.ready.sum$agecat[data.ready.sum$approximate_age >=65]                                          <- "65+"
data.ready.sum$agecat[data.ready.sum$approximate_age=="Missing"]                                    <- "Missing"
data.ready.sum$agecat[data.ready.sum$approximate_age == -9]                                         <- "Missing"

sum.age <- data.ready.sum %>%
  group_by(operator, weekpart, agecat) %>%
  summarise(freq = n(), weight = sum(weight), trip_weight = sum(trip_weight))

write.csv(sum.age, paste(D_OUTPUT_TABLEAU, "AgeCats.csv", sep = ""), row.names = FALSE, quote = T)

```
