---
title: "MTC Draft 2024 BART SAS Data Review"
output:
  html_document:
    theme: cosmo
    toc: true
---

```{r setup, echo=FALSE}
# Load necessary packages
library(tidyverse)
library(readxl)

# Load the data
USERPROFILE       <- gsub("////","/", Sys.getenv("USERPROFILE"))
Folder            <- file.path(USERPROFILE, "Box", "Modeling and Surveys")
Box_Surveys       <- file.path(Folder,"Surveys/Transit Passenger Surveys/Ongoing TPS/Individual Operator Efforts/BART 2024/BART MTC ETC RSG Project Folder/BART Pilot DATA - SAS/Renamed Versions")
other_stations_in <- file.path(Box_Surveys,"other_stations.xlsx")
oak_in            <- file.path(Box_Surveys,"oak.xlsx")
sfo_in            <- file.path(Box_Surveys,"sfo.xlsx")

other_stations    <- read_xlsx(other_stations_in,sheet = "BART ALL OTHER SAS Pilot Data")
oak               <- read_xlsx(oak_in, sheet = "OAK SAS PILOT DATA")
sfo               <- read_xlsx(sfo_in, sheet = "SFO SAS Pilot Data")


```

# Surveys by home county

```{r home_county, echo=FALSE}

household <- hh %>% 
  mutate(home_county = recode(home_county,
    "6001" = "Alameda",
    "6013" = "Contra Costa",
    "6041" = "Marin",
    "6055" = "Napa",
    "6075" = "San Francisco",
    "6081" = "San Mateo",
    "6085" = "Santa Clara",
    "6095" = "Solano",
    "6097" = "Sonoma"))

table(household$home_county)

household_joiner <- household %>% 
  select(hh_id,home_county)

```

# Trip characteristics

## Auto trip occupancy share - all trips

```{r all_trips_occupancy, echo=FALSE}
auto_trips <- trip %>%
  left_join(.,household_joiner,by="hh_id") %>% 
  mutate(num_travelers=recode(num_travelers,
                              "1"="1_Person",
                              "2"="2_Persons",
                              "3"="3_Persons",
                              "4"="4_Persons",
                              "5"="5p_Persons",
                              "995"="Missing")) %>% 
  filter(mode_type==8) 

all_auto_trips <- auto_trips %>% 
  group_by(home_county,num_travelers) %>% 
  summarize(occupancy=n()) %>% 
  mutate(share=round(100*occupancy/sum(occupancy),0)) %>% 
  select(-occupancy) %>% 
  pivot_wider(., names_from = num_travelers,values_from = share,values_fill = 0) %>% 
  select(-Missing)

bay_all_auto_trips <- auto_trips %>% 
  group_by(num_travelers) %>% 
  summarize(occupancy=n()) %>% 
  mutate(share=round(100*occupancy/sum(occupancy),0)) %>% 
  select(-occupancy) %>% 
  pivot_wider(., names_from = num_travelers,values_from = share,values_fill = 0) %>% 
  select(-Missing) %>% 
  mutate(home_county="Bay Area") %>% 
  relocate(home_county,.before = "1_Person")

all_auto_trips <- rbind(all_auto_trips,bay_all_auto_trips)

all_auto_trips
```

## Auto trip occupancy share - work trips

```{r work_trips_occupancy, echo=FALSE}

work_auto_trips <- auto_trips %>% 
  filter(d_purpose_category %in% c(2,3)) %>% 
  group_by(home_county,num_travelers) %>% 
  summarize(occupancy=n()) %>% 
  mutate(share=round(100*occupancy/sum(occupancy),0)) %>% 
  select(-occupancy) %>% 
  pivot_wider(., names_from = num_travelers,values_from = share,values_fill = 0) %>% 
  select(-Missing)

bay_work_auto_trips <- auto_trips %>% 
  filter(d_purpose_category %in% c(2,3)) %>% 
  group_by(num_travelers) %>% 
  summarize(occupancy=n()) %>% 
  mutate(share=round(100*occupancy/sum(occupancy),0)) %>% 
  select(-occupancy) %>% 
  pivot_wider(., names_from = num_travelers,values_from = share,values_fill = 0) %>% 
  select(-Missing) %>% 
  mutate(home_county="Bay Area") %>% 
  relocate(home_county,.before = "1_Person")

work_auto_trips <- rbind(work_auto_trips,bay_work_auto_trips)

work_auto_trips
```

# Demographics
## Compare race with ACS 2022 (Table C03002)

```{r compare_race, echo=FALSE}

race_variables <- c("Total"    = "C03002_001",
                    "White"    = "C03002_003",
                    "Black"    = "C03002_004",
                    "Other"    = "C03002_005",
                    "Asian"    = "C03002_006",
                    "Other"    = "C03002_007",
                    "Other"    = "C03002_008",
                    "Other"    = "C03002_009",
                    "Hispanic" = "C03002_012")

race <- get_acs(
  geography = "county",
  county = c("Alameda","Contra Costa","Marin","Napa","San Francisco","San Mateo","Santa Clara","Solano","Sonoma"),
  state = "california",
  variables = race_variables,
  year = 2022,
  survey = "acs1",
) %>% 
  group_by(NAME,variable) %>% 
  summarize(estimate=sum(estimate)) %>% 
  pivot_wider(.,names_from = variable,values_from = estimate) %>% 
  select(County=NAME,White,Black,Asian,Other,Hispanic,Total) %>% 
   mutate(County = gsub(" County, California", "", County))

race_bay <- get_acs(
  geography = "county",
  county = c("Alameda","Contra Costa","Marin","Napa","San Francisco","San Mateo","Santa Clara","Solano","Sonoma"),
  state = "california",
  variables = race_variables,
  year = 2022,
  survey = "acs1",
) %>% 
  group_by(variable) %>% 
  summarize(estimate=sum(estimate)) %>% 
  pivot_wider(.,names_from = variable,values_from = estimate) %>% 
  mutate(County="Bay Area") %>% 
  select(County,White,Black,Asian,Other,Hispanic, Total) 
  
race_all <- rbind(race,race_bay) %>% 
  mutate(
    ACS_White_Share    = round(100*White/Total,0),
    ACS_Black_Share    = round(100*Black/Total,0),
    ACS_Asian_Share    = round(100*Asian/Total,0),
    ACS_Other_Share    = round(100*Other/Total,0),
    ACS_Hispanic_Share = round(100*Hispanic/Total,0)) %>% 
  select(County,ACS_White_Share,ACS_Black_Share,ACS_Asian_Share,ACS_Other_Share,ACS_Hispanic_Share) %>% 
  mutate(ACS_Missing_Share=0) %>% 
  rename(Home_County=County)

# Now do survey

survey_race <- person %>% 
  filter(age>3) %>% 
  left_join(.,household_joiner,by="hh_id") %>% 
  mutate(race_recode=case_when(
    ethnicity_1==1 & race_5==1 & race_1==0 & race_2==0 & 
      race_3==0 & race_4==0 & race_997==0                                              ~ "White",
    ethnicity_1==1 & race_1==1 & race_5==0 & race_2==0 & 
      race_3==0 & race_4==0 & race_997==0                                              ~ "Black",
    ethnicity_1==1 & race_3==1 & race_5==0 & race_1==0 & race_2==0 & 
      race_4==0 & race_997==0                                                          ~ "Asian",
    ethnicity_1==1                                                                     ~ "Other",
    ethnicity_2==1 | ethnicity_3==1 | ethnicity_4==1 | ethnicity_997==1                ~ "Hispanic",
    ethnicity_999==1 | ethnicity_other==1                                              ~ "Other",
    ethnicity_1==995 & ethnicity_2==995 & ethnicity_3==995 & ethnicity_4==995 &
      ethnicity_997==995 & ethnicity_999==995 & race_1==995 & race_2==995 &
      race_3==995 & race_4==995 & race_5==995 & race_997==995 & race_999==995          ~ "Missing",
    TRUE                                                                               ~ "Miscoded")) 

survey_race_county <- survey_race %>% 
  group_by(home_county,race_recode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = race_recode,values_from = total) %>% 
  select(home_county,White,Black,Asian,Other,Hispanic,Missing)

survey_race_bay <- survey_race %>% 
  mutate(home_county="Bay Area") %>% 
  group_by(home_county,race_recode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = race_recode,values_from = total) %>% 
  select(home_county,White,Black,Asian,Other,Hispanic,Missing)

survey_race_all <- rbind(survey_race_county,survey_race_bay) %>% 
  mutate(Total=White+Black+Asian+Other+Hispanic+Missing,
         Survey_White_Share    = round(100*White/Total,0),
         Survey_Black_Share    = round(100*Black/Total,0),
         Survey_Asian_Share    = round(100*Asian/Total,0),
         Survey_Other_Share    = round(100*Other/Total,0),
         Survey_Hispanic_Share = round(100*Hispanic/Total,0),
         Survey_Missing_Share  = round(100*Missing/Total,0)) %>% 
  rename(Home_County=home_county)

combined_race <- left_join(race_all,survey_race_all,by="Home_County") %>% 
  select(Home_County,Survey_White_Share,ACS_White_Share,Survey_Black_Share,ACS_Black_Share,
         Survey_Asian_Share,ACS_Asian_Share,Survey_Other_Share,ACS_Other_Share,Survey_Hispanic_Share,
         ACS_Hispanic_Share,Survey_Missing_Share,ACS_Missing_Share)

```
### White

```{r white, echo=FALSE}

combined_race_white <- combined_race %>% 
  select(Home_County,Survey_White_Share,ACS_White_Share)

combined_race_white
```
### Black

```{r black, echo=FALSE}

combined_race_black <- combined_race %>% 
  select(Home_County,Survey_Black_Share,ACS_Black_Share)

combined_race_black
```
### Asian

```{r asian, echo=FALSE}

combined_race_asian <- combined_race %>% 
  select(Home_County,Survey_Asian_Share,ACS_Asian_Share)

combined_race_asian
```
### Other Race

```{r other, echo=FALSE}

combined_race_other <- combined_race %>% 
  select(Home_County,Survey_Other_Share,ACS_Other_Share)

combined_race_other
```
### Hispanic

```{r hispanic, echo=FALSE}

combined_race_hispanic <- combined_race %>% 
  select(Home_County,Survey_Hispanic_Share,ACS_Hispanic_Share)

combined_race_hispanic
```
### Missing Race

```{r missing_race, echo=FALSE}

combined_race_missing <- combined_race %>% 
  select(Home_County,Survey_Missing_Share,ACS_Missing_Share)

combined_race_missing
```
## Compare income with ACS 2022 (Table B19001)

```{r compare_income, echo=FALSE}

income_variables <- c("ACS_Total"         = "B19001_001",
                      "ACS_Under_25k"     = "B19001_002",
                      "ACS_Under_25k"     = "B19001_003",
                      "ACS_Under_25k"     = "B19001_004",
                      "ACS_Under_25k"     = "B19001_005",
                      "ACS_25k_to_50k"    = "B19001_006",
                      "ACS_25k_to_50k"    = "B19001_007",
                      "ACS_25k_to_50k"    = "B19001_008",
                      "ACS_25k_to_50k"    = "B19001_009",
                      "ACS_25k_to_50k"    = "B19001_010",
                      "ACS_50k_to_75k"    = "B19001_011",
                      "ACS_50k_to_75k"    = "B19001_012",
                      "ACS_75k_to_100k"   = "B19001_013",
                      "ACS_100k_to_200k"  = "B19001_014",
                      "ACS_100k_to_200k"  = "B19001_015",
                      "ACS_100k_to_200k"  = "B19001_016",
                      "ACS_Over_200k"     = "B19001_017")

income <- get_acs(
  geography = "county",
  county = c("Alameda","Contra Costa","Marin","Napa","San Francisco","San Mateo","Santa Clara","Solano","Sonoma"),
  state = "california",
  variables = income_variables,
  year = 2022,
  survey = "acs1",
) %>% 
  group_by(NAME,variable) %>% 
  summarize(estimate=sum(estimate)) %>% 
  pivot_wider(.,names_from = variable,values_from = estimate) %>% 
  select(County=NAME,ACS_Under_25k,ACS_25k_to_50k,ACS_50k_to_75k,ACS_75k_to_100k,ACS_100k_to_200k,ACS_Over_200k, ACS_Total) %>% 
   mutate(County = gsub(" County, California", "", County))

income_bay <- get_acs(
  geography = "county",
  county = c("Alameda","Contra Costa","Marin","Napa","San Francisco","San Mateo","Santa Clara","Solano","Sonoma"),
  state = "california",
  variables = income_variables,
  year = 2022,
  survey = "acs1",
) %>% 
  group_by(variable) %>% 
  summarize(estimate=sum(estimate)) %>% 
  pivot_wider(.,names_from = variable,values_from = estimate) %>% 
  mutate(County="Bay Area") %>% 
  select(County,ACS_Under_25k,ACS_25k_to_50k,ACS_50k_to_75k,ACS_75k_to_100k,ACS_100k_to_200k,ACS_Over_200k, ACS_Total) 

ACS_income_all <- rbind(income,income_bay) %>% 
  mutate(ACS_Missing=0,
         ACS_Under_25k_Share    =round(100*ACS_Under_25k/ACS_Total,0),
         ACS_25k_to_50k_Share   =round(100*ACS_25k_to_50k/ACS_Total,0),
         ACS_50k_to_75k_Share   =round(100*ACS_50k_to_75k/ACS_Total,0),
         ACS_75k_to_100k_Share  =round(100*ACS_75k_to_100k/ACS_Total,0),
         ACS_100k_to_200k_Share =round(100*ACS_100k_to_200k/ACS_Total,0),
         ACS_Over_200k_Share    =round(100*ACS_Over_200k/ACS_Total,0), 
         ACS_Missing_Share      =0)%>% 
  rename(Home_County=County)

survey_income <- household %>% 
  mutate(income_recode=case_when(
    income_broad==1                  ~ "Survey_Under_25k",
    income_broad==2                  ~ "Survey_25k_to_50k",
    income_broad==3                  ~ "Survey_50k_to_75k",
    income_broad==4                  ~ "Survey_75k_to_100k",
    income_broad==5                  ~ "Survey_100k_to_200k",
    income_broad==6                  ~ "Survey_Over_200k",
    income_broad %in% c(995,999)     ~ "Survey_Missing",
    TRUE                             ~ "Miscoded"
  ))


survey_income_county <- survey_income %>% 
  group_by(home_county,income_recode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = income_recode,values_from = total) %>% 
  select(home_county,Survey_Under_25k,Survey_25k_to_50k,Survey_50k_to_75k,Survey_75k_to_100k, 
    Survey_100k_to_200k,Survey_Over_200k,Survey_Missing) 
    

survey_income_bay <- survey_income %>% 
  mutate(home_county="Bay Area") %>% 
  group_by(home_county,income_recode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = income_recode,values_from = total) %>% 
  select(home_county,Survey_Under_25k,Survey_25k_to_50k,Survey_50k_to_75k,Survey_75k_to_100k, 
    Survey_100k_to_200k,Survey_Over_200k,Survey_Missing)

survey_income_all <- rbind(survey_income_county,survey_income_bay)%>% 
  mutate(Survey_Total=Survey_Under_25k + Survey_25k_to_50k + Survey_50k_to_75k + Survey_75k_to_100k + 
    Survey_100k_to_200k + Survey_Over_200k+Survey_Missing,
         Survey_Under_25k_Share    =round(100*Survey_Under_25k/Survey_Total,0),
         Survey_25k_to_50k_Share   =round(100*Survey_25k_to_50k/Survey_Total,0),
         Survey_50k_to_75k_Share   =round(100*Survey_50k_to_75k/Survey_Total,0),
         Survey_75k_to_100k_Share  =round(100*Survey_75k_to_100k/Survey_Total,0),
         Survey_100k_to_200k_Share =round(100*Survey_100k_to_200k/Survey_Total,0),
         Survey_Over_200k_Share    =round(100*Survey_Over_200k/Survey_Total,0), 
         Survey_Missing_Share      =round(100*Survey_Missing/Survey_Total,0))%>% 
  rename(Home_County=home_county)    

combined_income <- left_join(ACS_income_all,survey_income_all,by="Home_County") %>% 
  select(Home_County,Survey_Under_25k_Share,ACS_Under_25k_Share,Survey_25k_to_50k_Share,ACS_25k_to_50k_Share,
         Survey_50k_to_75k_Share,ACS_50k_to_75k_Share,Survey_75k_to_100k_Share,ACS_75k_to_100k_Share,
         Survey_100k_to_200k_Share,ACS_100k_to_200k_Share,Survey_Over_200k_Share,ACS_Over_200k_Share,
         Survey_Missing_Share, ACS_Missing_Share)

```
 
### Under_25k share

```{r under_25k, echo=FALSE}
combined_income_under_25k <- combined_income %>% 
  select(Home_County,Survey_Under_25k_Share,ACS_Under_25k_Share)

combined_income_under_25k

```
### 25_to_50k share

```{r 25_to_50k, echo=FALSE}
combined_income_25_to_50k <- combined_income %>% 
  select(Home_County,Survey_25k_to_50k_Share,ACS_25k_to_50k_Share)

combined_income_25_to_50k

```

### 50_to_75k share

```{r 50_to_75k, echo=FALSE}
combined_income_50_to_75k <- combined_income %>% 
  select(Home_County,Survey_50k_to_75k_Share,ACS_50k_to_75k_Share)

combined_income_50_to_75k

```

### 75_to_100k share

```{r 75_to_100k, echo=FALSE}
combined_income_75_to_100k <- combined_income %>% 
  select(Home_County,Survey_75k_to_100k_Share,ACS_75k_to_100k_Share)

combined_income_75_to_100k

```

### 100_to_200k share

```{r 100_to_200k, echo=FALSE}
combined_income_100_to_200k <- combined_income %>% 
  select(Home_County,Survey_100k_to_200k_Share,ACS_100k_to_200k_Share)

combined_income_100_to_200k

```

### Over 200k share

```{r 200kp, echo=FALSE}
combined_income_200kp <- combined_income %>% 
  select(Home_County,Survey_Over_200k_Share,ACS_Over_200k_Share)

combined_income_200kp

```

### Missing income share

```{r missing_income, echo=FALSE}
combined_income_missing <- combined_income %>% 
  select(Home_County,Survey_Missing_Share,ACS_Missing_Share)

combined_income_missing

```

# Compare rMove vs. browser vs. phone
## Distribution of survey method by county

```{r method_compare, echo=FALSE}

participation_joiner <- household %>% 
  mutate(participation_mode = case_when(
    participation_group %in% c(1,4,7)                ~ "browserMove",
    participation_group %in% c(2,5,8)                ~ "Call_Center",
    participation_group %in% c(3,6,9)                ~ "rMove",
    TRUE                                             ~ "Miscoded"
  )) %>% 
  select(hh_id,home_county,participation_mode,participation_group)

method_comparison <- participation_joiner %>% 
  group_by(home_county,participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>% 
  mutate(Total=Call_Center+browserMove+rMove)

method_comparison_bay <- participation_joiner %>% 
  group_by(participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>% 
  mutate(home_county="Bay Area",Total=Call_Center+browserMove+rMove)

method_comparison_all <- rbind(method_comparison,method_comparison_bay) %>% 
  rename(Home_County=home_county)

method_comparison_all

```

## Zero-trip days by data collection mode

```{r zero_trips, echo=FALSE}

zero_trip_days <- day %>% 
  left_join(.,participation_joiner,by="hh_id") %>% 
  filter(num_trips==0) %>% 
  group_by(home_county,participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>%
  rename(Call_Center0=Call_Center,browserMove0=browserMove,rMove0=rMove) %>% 
  mutate(Total0=Call_Center0+browserMove0+rMove0)

zero_trip_days_bay <- day %>% 
  left_join(.,participation_joiner,by="hh_id") %>% 
  filter(num_trips==0) %>% 
  group_by(participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>%
  rename(Call_Center0=Call_Center,browserMove0=browserMove,rMove0=rMove) %>% 
  mutate(home_county="Bay Area",Total0=Call_Center0+browserMove0+rMove0)

zero_trip_days_all <- rbind (zero_trip_days,zero_trip_days_bay)

num_days_mode <- day %>% 
  left_join(.,participation_joiner,by="hh_id") %>% 
  group_by(home_county,participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>% 
  mutate(Total=Call_Center+browserMove+rMove)

num_days_mode_bay <- day %>% 
  left_join(.,participation_joiner,by="hh_id") %>% 
  group_by(participation_mode) %>% 
  summarize(total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = total) %>% 
  mutate(home_county="Bay Area",Total=Call_Center+browserMove+rMove)

num_days_mode_all <- rbind(num_days_mode,num_days_mode_bay)

share_zero <- zero_trip_days_all %>% 
  left_join(.,num_days_mode_all,by="home_county") %>% 
  mutate(Call_Center_Share0=round(100*Call_Center0/Call_Center,0),
         browserMove_Share0=round(100*browserMove0/browserMove,0),
         rMove_Share0=round(100*rMove0/rMove,0),
         Total_Share0=round(100*Total0/Total,0)) %>% 
  rename(Home_County=home_county) %>% 
  select(Home_County,Call_Center_Share0,browserMove_Share0,rMove_Share0,Total_Share0)

share_zero

```

## Age of householder (Person 1) by collection mode

```{r participation_age, echo=FALSE}

person_joiner <- person %>% 
  select(hh_id,person_num,age) %>% 
  filter(person_num==1)

age_group <- participation_joiner %>% 
  left_join(.,person_joiner,by="hh_id") %>% 
  mutate(age_recode=case_when(
    age==1                           ~"Under 5",
    age==2                           ~"Ages 5 to 15",
    age==3                           ~"Ages 16 to 17",
    age==4                           ~"Ages 18 to 24",
    age==5                           ~"Ages 25 to 34",
    age==6                           ~"Ages 35 to 44",
    age==7                           ~"Ages 45 to 54",
    age==8                           ~"Ages 55 to 64",
    age==9                           ~"Ages 65 to 74",
    age==10                          ~"Ages 75 to 84",
    age==11                          ~"Ages 85 and up")) %>% 
  group_by(age_recode,participation_mode) %>%
  summarize(Total=n()) %>% 
  pivot_wider(names_from = participation_mode,values_from = Total) %>% 
  mutate(Total=Call_Center+browserMove+rMove)

age_group

```
## Share collection mode by age 
```{r participation_age_share, echo=FALSE}
age_group_share <- age_group %>% 
  mutate(
         Call_Center_Share=round(100*Call_Center/Total,0),
         browserMove_Share=round(100*browserMove/Total,0),
         rMove_Share=round(100*rMove/Total,0)) %>% 
  select (age_recode,Call_Center_Share,browserMove_Share,rMove_Share)

age_group_share
```
