---
title: "Tri-Delta On Off Exploration"
author: "David Ory"
date: "Monday, August 11, 2014"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

## Status
very early days

## Purpose
A key question for the transit on-board survey is whether or not there is value in performing an on/off pre-survey for small operators.  To help answer this question, we collected a deep on/off sample (approximately 70 percent of all ons and offs) for the Tri-Delta system.  What we want to know is (a) is it possible to craft logical weights based on boarding-alighting patterns for low ridership systems and, (b) if so, when and where and for whom would expansion weights based on route + direction + time of day be meaningfully different than weights based on route + time-of-day + direction + boarding-alighting?  

## Outputs
1.  `TODO` In-line charts and ugly tables

## TODO
1. Everything

## Procedure
```{r}
library(knitr)
suppressMessages(library(dplyr))
library(stringr)
library(optimx)
library(reshape2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

### Data reads
```{r data_reads}
input.tablet <- read.table(file = "M:/Data/OnBoard/Data and Reports/Tri Delta/As CSV/ONBOARD NO POUND OR SINGLE QUOTE.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.onoff  <- read.table(file = "M:/Data/OnBoard/Data and Reports/Tri Delta/As CSV/ON2OFF NO POUND OR SINGLE QUOTE.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.stops <- read.table(file = "M:/Data/OnBoard/Data and Reports/Tri Delta/As CSV/RTD_ROUTE_STOPS_ALL_MODES_MAR_2 NO POUND OR SINGLE QUOTE.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

## Prepare datasets
```{r prepare-datasets}
tablet <- input.tablet %>%
  select(tablet_id = ID, 
         route_dir = ROUTE.CODE, 
         direction = DIRECTION, 
         time_period = TIME_PERIOD, 
         board_id = BOARDING_LOCATION_STOPID, 
         board_shp = BOARDING_LOCATION_STOPID_SHP, 
         alight_id = ALIGHTING_LOCATION_STOPID, 
         alight_shp = ALIGHTING_LOCATION_STOPID_SHP, 
         weight = UNLINKED_WGHT_FCTR)

onoff <- input.onoff %>%
  select(onoff_id = ETC_ID, 
         ROUTE_ON, 
         DIRECTION_ON, 
         ROUTE_OFF, 
         DIRECTION_OFF, 
         time_period = TIME.PERIOD, 
         board_id = BOARDING_STOPID_, 
         BOARDING_STOPID_SHP, 
         board_seq = BOARDING_STOPSEQUENCE, 
         alight_id = ALIGHTING_STOPID_, 
         ALIGHTING_STOPID_SHP, 
         alight_seq = ALIGHTING_STOPSEQUENCE, 
         MATCH_CONFIDENCE = matchConfidence)

stops <- input.stops %>%
  select(operator = CPT_AGENCY, 
         route = SCH_ROUTED, 
         direction = RTD_PATTER, 
         stop_seq = SCH_STOPPO, 
         stop_shp = CPT_STOPPO) %>%
  filter(operator == '3D') %>%
  select(-operator)

# Suppress the route patterns and loops
stops <- stops %>%
  group_by(route, direction, stop_shp) %>%
  summarise(stop_seq = min(stop_seq))

# Join by route, direction, time of day, board, alight
tablet <- tablet %>%
  mutate(route = as.numeric(str_sub(route_dir, 1, 3))) %>%
  mutate(board_id = as.numeric(board_id)) %>%
  mutate(alight_id = as.numeric(alight_id)) %>%
  mutate(time_period = ifelse(time_period == "LATE PM", "EVENING", time_period))

onoff <- onoff %>%
  mutate(route = ROUTE_ON) %>%
  mutate(direction = DIRECTION_ON) %>%
  mutate(time_period = ifelse(time_period == "AM1", "EARLY AM", time_period)) %>%
  mutate(time_period = ifelse(time_period == "AM2", "AM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "MID", "MIDDAY",   time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM1", "PM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM2", "EVENING",  time_period))

onoff_sum <- onoff %>%
  group_by(route, direction, time_period, board_id, alight_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet, onoff_sum, by = c("route", "direction", "time_period", "board_id", "alight_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)
```



```{r route-segments-explore}

# How many stops does each route + direction have?
onoff_stops <- onoff %>%
  filter(MATCH_CONFIDENCE == "100%") %>%
  group_by(route, direction) %>%
  summarise(max_board_seq = max(board_seq), max_alight_seq = max(alight_seq))

# Use the number of stops to create break points
segment_breaks <- onoff_stops %>%
  mutate(break_2_1 = round((max_board_seq + max_alight_seq)/4)) %>%
  mutate(break_3_1 = round((max_board_seq + max_alight_seq)/6)) %>%
  mutate(break_3_2 = break_3_1 * 2) %>%
  mutate(break_4_1 = round((max_board_seq + max_alight_seq)/8)) %>%
  mutate(break_4_2 = break_4_1 * 2) %>%
  mutate(break_4_3 = break_4_1 * 3)

# Use the break points to put the ons and offs into boarding and alighting segments (TODO: make more elegant, generalizable)
onoff_breaks <- onoff %>%
  filter(MATCH_CONFIDENCE == "100%")

onoff_breaks <- left_join(onoff_breaks, segment_breaks, by = c("route", "direction"))

onoff_breaks <- onoff_breaks %>%
  mutate(board_seg_2  = ifelse(board_seq  <  break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq <  break_2_1, 1, 2)) %>%
  
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  
  mutate(board_seg_4  = ifelse(board_seq  < break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  
  mutate(alight_seg_4 = ifelse(alight_seq < break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

table(onoff_breaks$board_seg_2)
table(onoff_breaks$board_seg_3)
table(onoff_breaks$board_seg_4)

table(onoff_breaks$alight_seg_2)
table(onoff_breaks$alight_seg_3)
table(onoff_breaks$alight_seg_4)

# Use the break points to put the tablet into boarding and alighting segments
stops_board <- stops %>%
  select(route, direction, board_seq = stop_seq, board_shp = stop_shp)

stops_alight <- stops %>%
  select(route, direction, alight_seq = stop_seq, alight_shp = stop_shp)

tablet_breaks <- left_join(tablet,        stops_board,  by = c("route", "direction", "board_shp"))
tablet_breaks <- left_join(tablet_breaks, stops_alight, by = c("route", "direction", "alight_shp"))

tablet_breaks <- left_join(tablet_breaks, segment_breaks, by = c("route", "direction"))

tablet_breaks <- tablet_breaks %>%
  mutate(board_seg_2  = ifelse(board_seq  <  break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq <  break_2_1, 1, 2)) %>%
  
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  
  mutate(board_seg_4  = ifelse(board_seq  < break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  
  mutate(alight_seg_4 = ifelse(alight_seq < break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

table(tablet_breaks$board_seg_2)
table(tablet_breaks$board_seg_3)
table(tablet_breaks$board_seg_4)

table(tablet_breaks$alight_seg_2)
table(tablet_breaks$alight_seg_3)
table(tablet_breaks$alight_seg_4)


```


```{r check-matches}

# Step 1: join by route, access success
onoff_sum <- onoff %>%
  group_by(route) %>%
  summarise(onoff_count = n())
  
tab_onoff <- left_join(tablet, onoff_sum, by = c("route"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

route_missing <- tab_onoff %>%
  filter(onoff_count == 0)

table(route_missing$route)
table(onoff_sum$route)

# Results from Step 1: we lose 278 of 1259 because the on-off is missing routes 392, 393, 394, and 395

# Step 2: join non-missing routes, from Step 1, and direction
onoff_sum <- onoff %>%
  group_by(route, direction) %>%
  summarise(onoff_count = n())

tablet_route <- tablet %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395)
tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 2: once we account for routes, we do not lose anything else by adding direction

# Step 3: join non-missing routes, from Step 1, and direction, and time-of-day
onoff_sum <- onoff %>%
  group_by(route, direction, time_period) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "time_period"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 3: once we account for routes, we do not lose anything else by adding direction

# Step 4: join non-missing routes, from Step 1, and direction, time-of-day, and boarding location
onoff_sum <- onoff %>%
  group_by(route, direction, time_period, board_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "time_period", "board_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 4: once we account for routes, we lose 817 out of 981

# Step 5: join non-missing routes, from Step 1, and direction, boarding location (back off time from Step 4)
onoff_sum <- onoff %>%
  group_by(route, direction, board_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "board_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 5: once we account for routes and relax time of day, we lose 779 out of 981 (seems very high)

# Step 6: join non-missing routes, from Step 1, and direction, time-of-day, and boarding location
onoff_sum <- onoff %>%
  group_by(route, direction, time_period, alight_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "time_period", "alight_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 6: once we account for routes, we lose 832 out of 981

# Step 7: join non-missing routes, from Step 1, and direction, boarding location (back off time from Step 4)
onoff_sum <- onoff %>%
  group_by(route, direction, alight_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "alight_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 7: once we account for routes and relax time of day, we lose 779 out of 981 (seems very high)

# Step 8: join non-missing routes, from Step 1, and direction, time-of-day, boarding location, and alighting location
onoff_sum <- onoff %>%
  group_by(route, direction, time_period, board_id, alight_id) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_route, onoff_sum, by = c("route", "direction", "time_period", "board_id", "alight_id"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 8: we lose 964 out of 981

# Step 9: join non-missing routes, from Step 1, and direction, time-of-day, and two-stage boarding sequence
onoff_sum <- onoff_breaks %>%
  group_by(route, direction, time_period, board_seg_2) %>%
  summarise(onoff_count = n())

tablet_breaks_route <- tablet_breaks %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395)

tab_onoff <- left_join(tablet_breaks_route, onoff_sum, by = c("route", "direction", "time_period", "board_seg_2"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 9: we lose 109 out of 981

# Step 10: join non-missing routes, from Step 1, and direction, time-of-day, and three-stage boarding sequence
onoff_sum <- onoff_breaks %>%
  group_by(route, direction, time_period, board_seg_3) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_breaks_route, onoff_sum, by = c("route", "direction", "time_period", "board_seg_3"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 10: we lose 121 out of 981

# Step 11: join non-missing routes, from Step 1, and direction, time-of-day, and four-stage boarding sequence
onoff_sum <- onoff_breaks %>%
  group_by(route, direction, time_period, board_seg_4) %>%
  summarise(onoff_count = n())

tab_onoff <- left_join(tablet_breaks_route, onoff_sum, by = c("route", "direction", "time_period", "board_seg_4"))

tab_onoff <- tab_onoff %>%
  mutate(onoff_count = ifelse(is.na(onoff_count), 0, onoff_count))

table(tab_onoff$onoff_count)

# Results from Step 11: we lose 126 out of 981

# Okay: this seems to work well

```

```{r mini-clean}

# Mini-example, optimization, pass 3

# Consider a survey with five passengers on two routes
survey_id = c(1, 2, 3, 4, 5)
route = c(10, 10, 10, 10, 10)
direction = c("E", "E", "W", "W", "W")

survey_data = data.frame(survey_id, route, direction)

# Consider three targets: ridership on 10 E is 100, ridership on 10 W is 300, system ridership is 450
observed_targets_vector = c(target_1 = 100,
                            target_2 = 300,
                            target_3 = 450)

# Consider importance weights on the three targets of 1.00, 1.25, and 1.50
importance_weights_vector = c(weight_1 = 1.00,
                              weight_2 = 1.25,
                              weight_3 = 1.50)

# Create three incidence dummies on the survey for each target
survey_data <- survey_data %>%
  mutate(target_1_incidence = ifelse(route == 10 & direction == "E", 1, 0)) %>%
  mutate(target_2_incidence = ifelse(route == 10 & direction == "W", 1, 0)) %>%
  mutate(target_3_incidence = 1)

# Create unique weights to target incidence array
unique_weights <- survey_data %>%
  group_by(route, direction) %>%
  summarise(target_1_count = sum(target_1_incidence), 
            target_2_count = sum(target_2_incidence), 
            target_3_count = sum(target_3_incidence)) %>%
  ungroup() %>%
  select(target_1_count, target_2_count, target_3_count)

incidence_matrix <- data.matrix(unique_weights)

# Define the optimization function
optimization_function <- function(x, obs_target_v, import_v, inc_mtx) {
  
  # Compute estimated targets 
  est_target_v <- x %*% inc_mtx
  
  # Compute errors
  error_df <- data.frame(obs_target_v, est_target_v, import_v)
  error_df <- error_df %>%
    mutate(error <- import_v * abs(est_target_v - obs_target_v))
  
  # Return errors
  return(sum(error_df$error))

}

# Test the optimization function
x <- c(75, 100)
optimization_function(x, observed_targets_vector, importance_weights_vector, incidence_matrix)

# Find the optimal weights using OPTIMX
starting_weights_vector <- c(1,1)
optimal_weights <- optimx(starting_weights_vector, 
                          fn = optimization_function, 
                          obs_target_v = observed_targets_vector,
                          import_v = importance_weights_vector,
                          inc_mtx = incidence_matrix)


```

# Next steps:
## 1. Start in on a small example with the Tri-Delta data

```{r small-tri-delta-example}

# Step 1: build the observed targets vector from the on-off counts
obs_targets_df <- onoff %>%
  filter(MATCH_CONFIDENCE == "100%") %>%
  group_by(route, direction, time_period) %>%
  summarise(onoff_count = n()) %>%
  ungroup() %>%
  mutate(target_id = paste(route, "--", direction, "--", time_period, sep = "")) %>%
  select(target_id, onoff_count)

# Add in an all routes target, scaled up by 10 percent (to make the optimization do something)
all_routes_target_df <- data.frame(target_id = c("all_routes"), onoff_count = c(sum(obs_targets_df$onoff_count) * 1.10))

obs_targets_df <- rbind(obs_targets_df, all_routes_target_df)

observed_targets_vector <- obs_targets_df$onoff_count

# Step 2: build the importance weights vector (all 1s for now)
importance_weights_vector <- obs_targets_vector * 0 + 1

# Step 3: build the incidence matrix (summarised tablet rows * 99)
est_targets_df <- tablet %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395) %>%
  group_by(route, direction, time_period) %>%
  summarise(records = n())

est_targets_list_df <- est_targets_df %>%
  ungroup() %>%
  mutate(target_id = paste(route, "--", direction, "--", time_period, sep = "")) %>%
  select(target_id)

# Get all the target rows as columns
obs_targets_columns <- obs_targets_df %>%
  mutate(one = 1)
obs_targets_columns <- dcast(obs_targets_columns, target_id ~ target_id, value.var = "one")
obs_targets_columns[is.na(obs_targets_columns)] <- 0
obs_targets_columns <- obs_targets_columns %>%
  mutate(all_routes = 1)

# Join the target columns to the survey rows
est_targets_list_df <- left_join(est_targets_list_df, obs_targets_columns, by = c("target_id"))

est_targets_list_df <- est_targets_list_df %>%
  select(-target_id)

incidence_matrix <- data.matrix(est_targets_list_df)

# Step 4: check the optimization function
starting_weights_vector <- sample(1:length(est_targets_list_df$all_routes))
for (i in 1:length(starting_weights_vector)) starting_weights_vector[i] <- 1
x <- starting_weights_vector
optimization_function(x, observed_targets_vector, importance_weights_vector, incidence_matrix)

# Step 5: run the optimization
starting_weights_vector <- sample(1:length(est_targets_df$all_routes))
for (i in 1:length(starting_weights_vector)) starting_weights_vector[i] <- 1

optimal_weights <- optimx(starting_weights_vector, 
                          fn = optimization_function, 
                          obs_target_v = observed_targets_vector,
                          import_v = importance_weights_vector,
                          inc_mtx = incidence_matrix)

# Step 6: extract the weights
weights_df <- optimal_weights %>%
  filter(convcode == 0) %>%
  select(-convcode, -value, -fevals, -gevals, -niter, -kkt1, -kkt2, -xtimes)

weights_df <- as.data.frame(t(data.matrix(weights_df)))

tablet_weights <- cbind(est_targets_df, weights_df)

tablet_weights <- tablet_weights %>%
  select(route, direction, time_period, records, optimal_weights = V1)

# Step 7: compare the optimal weights to the ETC weights
tablet_etc_weights <- tablet %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395) %>%
  group_by(route, direction, time_period) %>%
  summarise(sum_weights = sum(weight))

# START HERE: join the above two files and inspect the differences









```

