---
title: "Build-Caltrain-Expansion-Inputs-Approach-Two"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
seems to be working; may need to iterate after expansion is worked out

#### TODO
1. Is it worth re-building the Category `B` targets from the source data?

#### Purpose
Build the `observed targets` and the `observed targets definitions` files for the Caltrain surveys.  The targets will be as follows:

1.  Category `A` - system ridership
2.  Category `B` - ridership by boarding station and alighting station
3.  Category `C` - ridership by line number
4.  Category `D` - ridership by service type (local, limited, baby bullet)

Two surveys were performed on Caltrain: an on-off survey and an personal interview survey.  The personal interview survey was then expanded in two separate ways: (a) by boarding, alighting, and time of day; and, (b) by route, time, and direction.

The Category `B` targets will be derived from the on/off weights applied in the survey rather than the original on/off source data. Caltrains counts (http://www.caltrain.com/about/statsandreports/Ridership.html) will be used for Categories `C` and `D`.  

#### Overhead
```{r}
library(knitr)
library(stringr)
suppressMessages(library(dplyr))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data Preparation
#### Remote data locations
```{r data-reads}
input_dir  <- "M:/Data/OnBoard/Data and Reports/Caltrain/As CSV/"
output_dir <- "M:/Data/OnBoard/Data and Reports/Caltrain/multi-criteria-expansion/"

F_SURVEY <- paste(input_dir, "Caltrain_Final_Submitted_1_5_2015_withSERVICETYPE NO POUND OR SINGLE QUOTE.csv", sep = "")
F_LINE <- paste(input_dir, "ridership-by-line-2014.csv", sep = "")

survey <- read.table(file = F_SURVEY, header = TRUE, sep = ",", stringsAsFactors = FALSE)
line_ridership <- read.table(file = F_LINE, header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

## Build the targets and definitions
#### Definitions
```{r definitions}
# Definitions
observed_targets_defn <- data.frame(target_category_id = c("A", "B", "B", "C", "D"), 
                                    survey_variable    = c("all_routes",
                                                           "board_station", "alight_station",
                                                           "line_number",
                                                           "service_type"))


```

#### Category A - system ridership
```{r a-targets}
# Check the different sums
sum(line_ridership$total_boardings)

survey <- survey %>%
  mutate(LINKED_WGT_FCTR_BtoA = ifelse(is.na(LINKED_WGT_FCTR_BtoA), 0, LINKED_WGT_FCTR_BtoA)) %>%
  mutate(LINKED_WGT_FCTR_RTD = ifelse(is.na(LINKED_WGT_FCTR_RTD), 0, LINKED_WGT_FCTR_RTD))

sum(survey$LINKED_WGT_FCTR_BtoA)
sum(survey$LINKED_WGT_FCTR_RTD)

# Use 43,400

# Counts
observed_targets <- data.frame(target_count       = c(43400), 
                               importance_weight  = c(30),
                               target_category_id = c("A"),
                               line_number        = c(NA),
                               board_station      = c(NA),
                               alight_station     = c(NA),
                               service_type       = c(NA))

```

#### Category B
```{r b-targets}
target_b <- survey %>%
  group_by(ENTER_STATION, EXIT_STATION) %>%
  summarise(target_count = sum(LINKED_WGT_FCTR_BtoA)) %>%
  ungroup() %>%
  mutate(target_category_id = "B") %>%
  mutate(importance_weight = 20) %>%
  mutate(line_number = NA) %>%
  rename(board_station  = ENTER_STATION) %>%
  rename(alight_station = EXIT_STATION) %>%
  mutate(service_type = NA)

# Bind with A
observed_targets <- rbind(observed_targets, target_b)

```

#### Category C
Scale boardings to match Target A to improve optimization efficiency
```{r c-targets}
target_c <- line_ridership %>%
  rename(line_number = train_number) %>%
  mutate(target_count = total_boardings * 43400.0 / 52615.0) %>%
  select(-total_boardings) %>%
  mutate(target_category_id = "C") %>%
  mutate(importance_weight = 1) %>%
  mutate(board_station = NA) %>%
  mutate(alight_station = NA) %>%
  mutate(service_type = NA)

# Bind with A+B
observed_targets <- rbind(observed_targets, target_c)

```

#### Category D
```{r d-targets}
route_service <- survey %>%
  select(train_number = RUN, service_type = SERVICE_TYPE) %>%
  group_by(train_number, service_type) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  select(-count)

target_d <- left_join(line_ridership, route_service, by = c("train_number"))

target_d <- target_d %>%
  group_by(service_type) %>%
  summarise(target_count = sum(total_boardings)) %>%
  ungroup() %>%
  filter(!is.na(service_type)) %>%
  mutate(target_count = target_count * 43400.0 / 52615.0) %>%
  mutate(target_category_id = "D") %>%
  mutate(importance_weight = 2) %>%
  mutate(line_number = NA) %>%
  mutate(board_station = NA) %>%
  mutate(alight_station = NA)

names(observed_targets)
names(target_d)

# Bind with A+B+C
observed_targets <- rbind(observed_targets, target_d)

# Add a unique integer
target_id <- seq(1:length(observed_targets$target_category_id))
observed_targets <- cbind(target_id, observed_targets)

```


## Data writes
```{r writes}
write.csv(observed_targets_defn, 
          file = paste(output_dir, "observed-target-definitions-two.csv", sep = ""), 
          row.names = FALSE, 
          quote = F)

save(observed_targets_defn, file = paste(output_dir, "observed-target-definitions-two.Rdata", sep = ""))

write.csv(observed_targets, 
          file = paste(output_dir, "observed-target-counts-two.csv", sep = ""), 
          row.names = FALSE, 
          quote = F)

save(observed_targets, file = paste(output_dir, "observed-target-counts-two.Rdata", sep = ""))


```
