---
title: "Build CHAMP Crosswalk"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes a Muni route segmentation file and a SF CHAMP flow file.  It attempts to build a crosswalk relating the stop segment to each of the stops in the CHAMP file.  

#### Outputs
1.  A database relating the CHAMP stops to the Muni segments

#### _TODO_
1. Everything; early days

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(reshape2)
library(stringr)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O
```{r remote-io}
F_CHAMP_FLOWS = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/champ-flows/Muni_OD_no_pound.csv"
F_MUNI_SEGS =   "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/MuniLineSegmentation_2015-09-29.csv"

F_OUTPUT = "muni-champ-segments-database-automated.csv"

```

#### Parameters
```{r champ-rename}
champ_rename <- data.frame(champ_name = c("14L", "14SHORT", "1SHORT", "24EA", "28L", "30SHORT", "38_33RD", "38FTM",
                                          "38L", "5EA",     "5SHORT", "71",   "71L", "80X",     "8X",      "94L",
                                          "9L",  "K"),
                           route_man  = c("14R", "14",     "1",       "24",   "28R", "30",      "38",      "38",
                                          "38R", "5",      "5",       "7",    "7R",  "DUNO",    "8AX",     "DUNO",
                                          "9R",  "KT"))

champ_rename <- champ_rename %>%
  mutate(route_man = paste(route_man))

```


#### Prepare the Muni Segmentation data
```{r prep-muni}
# Read in
segments_input <- read.table(file = F_MUNI_SEGS, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Melt
segments_df <- melt(segments_input, id.vars = "Route")

# Clean up
segments_df <- segments_df %>%
  rename(location = value) %>%
  mutate(variable = paste(variable))

# Columns from the variable column
variable_split <- read.table(text = segments_df$variable, sep = ".", colClasses = "character")
variable_split <- variable_split %>%
  rename(segment_number = V2, start_or_end = V3) %>%
  select(segment_number, start_or_end)
  
# More clean up
segments_df <- segments_df %>%
  select(route = Route, location)

segments_df <- cbind(segments_df, variable_split)

segments_df <- arrange(segments_df, route, segment_number, desc(start_or_end))

segments_df <- segments_df %>%
  filter(nchar(location) > 0)

remove(variable_split, segments_input)

```

#### Prepare the CHAMP data
```{r prep-champ}
champ_input <- read.table(file = F_CHAMP_FLOWS, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Extract the Muni name
champ_df <- champ_input %>%
  mutate(dir = str_sub(champ_name, start = -1)) %>%
  mutate(direction = "MISSING") %>%
  mutate(direction = ifelse(dir == "I", "INBOUND", direction)) %>%
  mutate(direction = ifelse(dir == "O", "OUTBOUND", direction)) %>%
  select(-dir) %>%
  filter(direction != "MISSING") %>%
  mutate(route = str_sub(champ_name, start = 4)) %>%
  mutate(route = str_sub(route, start = 1, end = -2))

table(champ_df$route)
table(champ_df$direction)

# START HERE:
# 1. get the below join to work
# 2. make sure most of the routes join
# 3. move on to figuring out the problem: how to build the output

# Add in the manual names
champ_df <- left_join(champ_df, champ_rename, by = c("route"))

# Check the join with the segment data
champ_join <- champ_df %>%
  group_by(route) %>%
  summarise(champ_count = n())

segments_join <- segments_df %>%
  group_by(route) %>%
  summarise(segments_count = n())

table(champ_join$route)
table(segments_join$route)

join_test <- left_join(champ_join, segments_join, by = c("route"))

join_errors <- join_test %>%
  filter(is.na(segments_count))

table(join_errors$route)





```

