---
title: "Build CHAMP Crosswalk"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes a Muni route segmentation file and a SF CHAMP flow file.  It attempts to build a crosswalk relating the stop segment to each of the stops in the CHAMP file.  

#### Outputs
1.  A database relating the CHAMP stops to the Muni segments

#### _TODO_
1. The matching was not as successful as I'd hoped, so I did not spend much time refining the sequence adjustment.  Seems where I'm at now may provide 90 percent of the ultimate boost, with the additional 10 percent likely not worth the effort.
2.  Ask Drew about 80X and 94L: what Muni routes do these belong to

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(reshape2)
library(stringr)
library(stringdist)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O
```{r remote-io}
F_CHAMP_FLOWS = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/champ-flows/Muni_OD_no_pound.csv"
F_MUNI_SEGS =   "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/MuniLineSegmentation_2015-09-29.csv"

F_OUTPUT = "muni-champ-segments-database-automated.csv"
F_OUTPUT_MATCHES = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/champ-segments-matches.csv"

```

#### Parameters
```{r champ-rename}
champ_rename <- data.frame(route = c("14L", "14SHORT", "1SHORT", "24EA", "28L", "30SHORT", "38_33RD", "38FTM",
                                          "38L", "5EA",     "5SHORT", "71",   "71L", "80X",     "8X",      "94L",
                                          "9L",  "K"),
                           route_man  = c("14R", "14",     "1",       "24",   "28R", "30",      "38",      "38",
                                          "38R", "5",      "5",       "7",    "7R",  "DUNO",    "8AX",     "DUNO",
                                          "9R",  "KT"))

champ_rename <- champ_rename %>%
  mutate(route = paste(route)) %>%
  mutate(route_man  = paste(route_man))

```


#### Prepare the Muni Segmentation data
```{r prep-muni}
# Read in
segments_input <- read.table(file = F_MUNI_SEGS, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Melt
segments_df <- melt(segments_input, id.vars = "Route")

# Clean up
segments_df <- segments_df %>%
  rename(location = value) %>%
  mutate(variable = paste(variable))

# Columns from the variable column
variable_split <- read.table(text = segments_df$variable, sep = ".", colClasses = "character")
variable_split <- variable_split %>%
  rename(segment_number = V2, start_or_end = V3) %>%
  select(segment_number, start_or_end)
  
# More clean up
segments_df <- segments_df %>%
  select(route = Route, location)

segments_df <- cbind(segments_df, variable_split)

segments_df <- arrange(segments_df, route, segment_number, desc(start_or_end))

segments_df <- segments_df %>%
  filter(nchar(location) > 0)

remove(variable_split, segments_input)

```

#### Prepare the CHAMP data
```{r prep-champ}
champ_input <- read.table(file = F_CHAMP_FLOWS, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Extract the Muni name
champ_df <- champ_input %>%
  mutate(dir = str_sub(champ_name, start = -1)) %>%
  mutate(direction = "MISSING") %>%
  mutate(direction = ifelse(dir == "I", "INBOUND", direction)) %>%
  mutate(direction = ifelse(dir == "O", "OUTBOUND", direction)) %>%
  select(-dir) %>%
  filter(direction != "MISSING") %>%
  mutate(route = str_sub(champ_name, start = 4)) %>%
  mutate(route = str_sub(route, start = 1, end = -2))

table(champ_df$route)
table(champ_df$direction)

# Add in the manual names
champ_df <- left_join(champ_df, champ_rename, by = c("route"))

champ_df <- champ_df %>%
  mutate(route = ifelse(is.na(route_man), route, route_man)) %>%
  select(-route_man)

table(champ_df$route)

# Check the join with the segment data
champ_join <- champ_df %>%
  group_by(route) %>%
  summarise(champ_count = n())

segments_join <- segments_df %>%
  group_by(route) %>%
  summarise(segments_count = n())

table(champ_join$route)
table(segments_join$route)

join_test <- left_join(champ_join, segments_join, by = c("route"))

join_errors <- join_test %>%
  filter(is.na(segments_count))

table(join_errors$route)

remove(champ_join, segments_join, join_test, champ_input)

```

#### String matching
```{r string-matching}
# condense the champ flows
champ_work <- champ_df %>%
  group_by(champ_name, route, direction, SeqB) %>%
  summarise(StreetsB = max(StreetsB))

# initialize the output data frame
outcome_df <- data.frame (index = integer(),
                          champ_name = character(),
                          route = character(),
                          direction = character(),
                          SeqB = integer(),
                          StreetsB = character(),
                          location = character(),
                          segment_number = character(),
                          start_or_end = character())

# loop through each route
for (rte in unique(segments_df$route)){
  
  print(rte)
  
  seg_rte_df <- segments_df %>%
    filter(route == rte)
  
  chp_rte_df <- champ_work %>%
    filter(route == rte)
  
  # loop through each direction
  for (dir in unique(chp_rte_df$direction)){
    
    chp_rte_dir_df <- chp_rte_df %>%
      filter(direction == dir)
    
    # manually change Army to Cesar Chavez
    chp_rte_dir_df <- chp_rte_dir_df %>%
      mutate(StreetsB = str_replace(StreetsB, "Army", "Cesar Chavez"))
    
    # separate the champ location street names (start with five)
    streets_df <- colsplit(chp_rte_dir_df$StreetsB, "/", c("street_1", "street_2", "street_3", "street_4", "street_5"))
    
    chp_rte_dir_df <- cbind(chp_rte_dir_df, streets_df)
    
  
    
    # loop through each location
    for (loc in unique(seg_rte_df$location)){
      
      seg_rte_loc_df <- seg_rte_df %>%
        filter(location == loc)
      
      # separate the key location into pieces (breaking first on &, then on /) (start with six total bins)
      streets_df <- colsplit(seg_rte_loc_df$location, "&", c("street_a", "street_b"))
      streets_a  <- colsplit(streets_df$street_a, "/", c("target_street_1", "target_street_2", "target_street_3"))
      streets_b  <- colsplit(streets_df$street_b, "/", c("target_street_4", "target_street_5", "target_street_6"))
      
      seg_rte_loc_df <- cbind(seg_rte_loc_df, streets_a, streets_b)
      
      # join the location to the champ flows
      string_match <- left_join(chp_rte_dir_df, seg_rte_loc_df, by = c("route"))
      
      # flatten the join to get all the street and target street side-by-side
      string_match_flat <- melt(string_match, 
                                id = c("champ_name", "route", "direction", "SeqB", 
                                       "StreetsB", "location", "segment_number", "start_or_end", 
                                       "street_1", "street_2", "street_3", "street_4", "street_5"), 
                                na.rm = TRUE)
      
      # compare each source street against each target
      string_match_flat <- string_match_flat %>%
        select(-variable) %>%
        rename(target_street = value) %>%
        mutate(score_1 = ifelse(is.na(street_1) | str_length(street_1) < 2 | str_length(target_street) < 2,
                                50, stringdist(target_street, street_1, method = "dl"))) %>%
        mutate(score_2 = ifelse(is.na(street_2) | str_length(street_2) < 2 | str_length(target_street) < 2,
                                50, stringdist(target_street, street_2, method = "dl"))) %>%
        mutate(score_3 = ifelse(is.na(street_3) | str_length(street_3) < 2 | str_length(target_street) < 2,
                                50, stringdist(target_street, street_3, method = "dl"))) %>%
        mutate(score_4 = ifelse(is.na(street_4) | str_length(street_4) < 2 | str_length(target_street) < 2,
                                50, stringdist(target_street, street_4, method = "dl"))) %>%
        mutate(score_5 = ifelse(is.na(street_5) | str_length(street_5) < 2 | str_length(target_street) < 2,
                                50, stringdist(target_street, street_5, method = "dl"))) %>%
        mutate(aggregate_score = pmin(score_1, score_2, score_3, score_4, score_5))
      
      # condense the flatten file back up
      string_match_scored <- string_match_flat %>%
        group_by(champ_name, route, direction, SeqB, StreetsB, location, segment_number, start_or_end) %>%
        summarise(aggregate_score = sum(aggregate_score))
      
      # keep the winner
      keeper <- string_match_scored %>%
        ungroup() %>%
        filter(aggregate_score == min(string_match_scored$aggregate_score))
      
      outcome_df <- rbind(outcome_df, keeper)
      
    } # loc
    
  } # dir
  
} # rte

# update the index
rownames(outcome_df) <- 1:nrow(outcome_df)

# pick a winner among the ties
outcome_ties_df <- outcome_df %>%
  group_by(champ_name, direction, route, location, segment_number, start_or_end, aggregate_score) %>%
  summarise(SeqB = min(SeqB))

# put StreetsB back on
outcome_ties_df <- left_join(outcome_ties_df, outcome_df, by = c("champ_name", "direction", "route", "segment_number", "start_or_end", "location", "aggregate_score", "SeqB"))

outcome_ties_df <- outcome_ties_df %>%
  mutate(confidence = "high") %>%
  mutate(confidence = ifelse(aggregate_score > 2, "medium", confidence)) %>%
  mutate(confidence = ifelse(aggregate_score > 10, "low", confidence))

outcome_ties_df <- arrange(outcome_ties_df, aggregate_score)

remove(champ_work, chp_rte_df, chp_rte_dir_df, seg_rte_df, seg_rte_loc_df, streets_a, streets_b, streets_df,
       string_match, string_match_flat, string_match_scored, outcome_df, keeper, join_errors)

```

#### Read and write matches to allow for manual manipulation
```{r write-matches}
write.csv(outcome_ties_df, file = F_OUTPUT_MATCHES, row.names = FALSE, quote = F)

```

```{r read-matches}
matches_df <- read.table(file = F_OUTPUT_MATCHES, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# for now, use results with a matching score of 5 or lower
matches_df <- matches_df %>%
  filter(aggregate_score < 6)

remove(outcome_ties_df)

```

#### Segment CHAMP data
```{r segment-champ}
# Step 1: Build the output data I want without the segment labels
# Use the 'B' location from each link, as in from A to B
champ_segments_df <- champ_df %>%
  select(champ_name, apc_route = route, apc_direction = direction, stop_sequence = SeqB, stop_location = StreetsB) %>%
  group_by(champ_name, apc_route, apc_direction, stop_sequence) %>%
  summarise(stop_location = max(stop_location))

# Since we used the 'B', we are missing the first stops, add them here
first_df <- champ_df %>%
  select(champ_name, apc_route = route, apc_direction = direction, stop_sequence = SeqA, stop_location = StreetsA) %>%
  filter(stop_sequence == 0) %>%
  group_by(champ_name, apc_route, apc_direction, stop_sequence) %>%
  summarise(stop_location = max(stop_location))

champ_segments_df <- rbind(champ_segments_df, first_df)
champ_segments_df <- arrange(champ_segments_df, apc_route, apc_direction, champ_name, stop_sequence, stop_location)
champ_segments_df <- ungroup(champ_segments_df)

# Step 2: join the segment labels
outcome_join <- matches_df %>%
  select(champ_name, 
         apc_route = route, 
         apc_direction = direction, 
         stop_sequence = SeqB, 
         stop_location = StreetsB, 
         match_location = location,
         segment_number,
         start_or_end)

champ_segments_df <- left_join(champ_segments_df, outcome_join, by = c("champ_name", 
                                                                       "apc_route", 
                                                                       "apc_direction", 
                                                                       "stop_sequence", 
                                                                       "stop_location"))

# Step 3: fill in the gaps
route_vector <- champ_segments_df$apc_route
dir_vector <- champ_segments_df$apc_direction
segment_vector <- champ_segments_df$segment_number
updated_segment_number <- vector(mode="character", length = length(route_vector))

rolling_segment = "Missing"
prior_route = "Missing"
prior_direction = "Missing"
for (i in 1:length(route_vector)) {
  
  # check if route and direction are the same, carry over the roller
  if (route_vector[i] == prior_route & dir_vector[i] == prior_direction) {
    
    if (is.na(segment_vector[i])){
      updated_segment_number[i] = rolling_segment
    } else {
      updated_segment_number[i] = segment_vector[i]
      rolling_segment = segment_vector[i]
    }
    
    
  } else {
    
    if (is.na(segment_vector[i])) updated_segment_number[i] = "Missing"
    else updated_segment_number[i] = segment_vector[i]
    
    rolling_segment = updated_segment_number[i]
    
  }
  
  # update the running variables
  prior_route = route_vector[i]
  prior_direction = dir_vector[i]

}

champ_segments_df <- cbind(champ_segments_df, updated_segment_number)

  
```

#### Write out the working file
```{r write}
write.csv(champ_segments_df, file = F_OUTPUT, row.names = FALSE, quote = F)

```

