---
title: "Muni Travel Model Priors"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes boarding and alighting data from SF Muni automated passenger counters, boarding and alighting flows from the SFCTA SF-CHAMP travel model, route segment definitions provided by SF Muni, and a crosswalk connecting the APC and travel model data.  It combines this data to generate an estimate of boarding and alighting flows consistent with both the APC and travel model data for potenital use as prior estimates of flows on SF Muni.  

#### Outputs
1.  A consolidated database of flow estimates
2.  A summary database comparing boardings and alightings across the sources

#### _TODO_
1.  Do we have all the routes? 

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(reshape2)
suppressMessages(library(dplyr))
library(stringr)
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Paramaters
```{r parameters}
# APC time codes: am (6 to 9 am), mid (9 am to 2 pm), schl (2 to 4 pm), pm (4 to 7 pm), eve (7 to 10 pm), night (10 pm to 1 am), owl (1 to 6 am), daily (all)
# CHAMP time codes: EA (3 to 6 am), AM (6 to 9 am), MD (9 am to 3:30 pm), PM (3:30 to 6:30 pm), EV (6:30 pm to 3 am)

champ_time_of_day = c("AM", "MD",  "MD",   "PM", "EV",  "EV",    "EV",  "daily")
apc_time_of_day   = c("am", "mid", "schl", "pm", "eve", "night", "owl", "daily")
time_of_day_df <- data.frame(apc_time_of_day, champ_time_of_day)

# Observed rail ridership
#             F In  F Out  J in J out   K in  K out   L in  L out   M in  M out   N in  N out
OBS_RAIL = c(11304, 11304, 8157, 8157, 17763, 17763, 15488, 15488, 14326, 14326, 22166, 22166) 

# Champ rail names
CHAMP_RAIL_NAMES = c("MUNFI", "MUNFO", "MUNJI", "MUNJO", "MUNKI", "MUNKO", 
                     "MUNLI", "MUNLO", "MUNMI", "MUNMO", "MUNNI", "MUNNO")

# Log linear fit parameters
THRESH = 1.e-07
ITER = 1000

```

#### Remote I/O Locations
```{r remote-io}
F_APC_COUNTS     = "M:/Data/Transit/Muni APC Through Time/consolidated-database.csv"
F_CHAMP_FLOWS    = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/champ-flows/Muni_OD_no_pound.csv"

# F_APC_COUNTS     = "~DavidWork/Documents/consolidated-database.csv"
# F_CHAMP_FLOWS    = "~DavidWork/Documents/Muni_OD_no_pound.csv"

F_APC_SEGMENTS   = "muni-apc-segments-database.csv"
F_CHAMP_SEGMENTS = "muni-champ-segments-database.csv"

F_OUTPUT_PRIORS = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/priors.csv"
F_OUTPUT_COMPS  = "M:/Data/OnBoard/Data and Reports/Muni/travel-model-priors/priors-comparisons.csv"

# F_OUTPUT_PRIORS = "~DavidWork/Documents/test-output-priors.csv"
# F_OUTPUT_COMPS  = "~DavidWork/Documents/test-output-comparisons.csv"
```

#### Data reads
```{r data-reads}
apc_counts     <- read.table(file = F_APC_COUNTS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
champ_flows    <- read.table(file = F_CHAMP_FLOWS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
apc_segments   <- read.table(file = F_APC_SEGMENTS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
champ_segments <- read.table(file = F_CHAMP_SEGMENTS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

#### Prepare CHAMP data
```{r prep-champ}
champ_df <- champ_flows %>%
  select(champ_name, board_stop_sequence = SeqA, alight_stop_sequence = SeqB, time_of_day = TimeOfDay, flow = Trips)

# First join to get APC route name and direction
join_A <- champ_segments %>%
  group_by(champ_name, apc_route, apc_direction) %>%
  summarise(count = n()) %>%
  select(-count)
champ_df <- left_join(champ_df, join_A, by = c("champ_name"))

# Next join the boarding segment
join_B <- champ_segments %>%
  select(champ_name, 
         champ_board_stop_location = stop_location, 
         board_stop_sequence = stop_sequence, 
         board_segment = segment)
champ_df <- left_join(champ_df, join_B, by = c("champ_name", "board_stop_sequence"))

# Next join the alighting segment
join_C <- champ_segments %>%
  select(champ_name, 
         champ_alight_stop_location = stop_location, 
         alight_stop_sequence = stop_sequence, 
         alight_segment = segment)
champ_df <- left_join(champ_df, join_C, by = c("champ_name", "alight_stop_sequence"))

# Summarize
champ_sum <- champ_df %>%
  group_by(champ_name, apc_route, apc_direction, time_of_day, board_segment, alight_segment) %>%
  summarise(flow = sum(flow))

# Add in the daily sums
daily <- champ_sum %>%
  group_by(champ_name, apc_route, apc_direction, board_segment, alight_segment) %>%
  summarise(flow = sum(flow)) %>%
  ungroup() %>%
  mutate(time_of_day = "daily")

champ_sum <- rbind(champ_sum, daily)

remove(join_A, join_B, join_C, champ_df, daily)

```

#### Extract rail counts from CHAMP data
```{r rail-counts-from-champ}
# Build a crosswalk
champ_name = CHAMP_RAIL_NAMES
obs_riders <- OBS_RAIL  
rail_names <- data.frame(champ_name, obs_riders)
rail_names <- rail_names %>%
  mutate(is_rail = TRUE)

# Adjust the ridership to observed
to_adjust <- champ_sum %>%
  group_by(champ_name) %>%
  filter(!(time_of_day == "daily")) %>%
  summarise(flow = sum(flow))

to_adjust <- left_join(to_adjust, rail_names, by = c("champ_name"))
to_adjust <- na.omit(to_adjust)
to_adjust <- to_adjust %>%
  mutate(adj_fact = obs_riders / flow) %>%
  select(champ_name, adj_fact)

champ_sum <- left_join(champ_sum, to_adjust, by = c("champ_name"))
champ_sum <- champ_sum %>%
  mutate(flow = ifelse(is.na(adj_fact), flow, flow * adj_fact)) %>%
  select(-adj_fact)

# Extract the data
rail_df <- left_join(champ_sum, rail_names, by = c("champ_name"))
rail_df <- na.omit(rail_df)
rail_df <- rail_df %>%
  select(-obs_riders)

# Boardings, then alightings
rail_board <- rail_df %>%
  select(apc_route, apc_direction, segment = board_segment, time_of_day, boardings = flow) %>%
  group_by(apc_route, apc_direction, segment, time_of_day) %>%
  summarise(boardings = sum(boardings))

rail_alight <- rail_df %>%
  select(apc_route, apc_direction, segment = alight_segment, time_of_day, alightings = flow) %>%
  group_by(apc_route, apc_direction, segment, time_of_day) %>%
  summarise(alightings = sum(alightings))

# Join
rail_counts <- merge(rail_board, rail_alight, by = c("apc_route", "apc_direction", "time_of_day", "segment"), all = TRUE)
rail_counts <- rail_counts %>%
  rename(route = apc_route, direction = apc_direction)

remove(rail_names, champ_name, obs_riders, to_adjust, rail_df, rail_board, rail_alight)

```

#### Prepare APC data 
```{r prep-apc}
# Select one tranche of counts and the data I need
apc_df <- apc_counts %>%
  filter(start_date == "2015-02-02" | start_date == "2/2/2015") %>%
  filter(week_part == "WEEKDAYS") %>%
  select(route, direction, stop_location, stop_sequence, time_of_day, boardings, alightings)

# Join the segments and filter
apc_df <- left_join(apc_df, apc_segments, by = c("route", "direction", "stop_sequence"))
apc_df <- na.omit(apc_df)

# Put in champ time-of-day
to_champ_time <- time_of_day_df %>%
  select(time_of_day = apc_time_of_day, champ_time_of_day)

apc_df <- left_join(apc_df, to_champ_time, by = c("time_of_day"))

apc_df <- apc_df %>%
  select(-time_of_day) %>%
  rename(time_of_day = champ_time_of_day)

# Sum by segment
apc_sum <- apc_df %>%
  group_by(route, direction, time_of_day, segment) %>%
  summarise(boardings = sum(boardings), alightings = sum(alightings))

table(apc_sum$time_of_day)

# Join with the rail data
apc_sum <- rbind(apc_sum, rail_counts)

remove(to_champ_time, apc_df)

```

#### Get segment names from the crosswalks
```{r segment-names}
# join the segments
champ_name = CHAMP_RAIL_NAMES
champ_rail_segments <- champ_segments[which(champ_segments[,1] %in% champ_name),]

champ_rail_segments <- champ_rail_segments %>%
  select(route = apc_route, direction = apc_direction, stop_location, stop_sequence, segment)

both_segments <- rbind(apc_segments, champ_rail_segments)

segment_names <- both_segments %>%
  group_by(route, direction, segment) %>%
  summarise(first_stop_sequence = min(stop_sequence), last_stop_sequence = max(stop_sequence))

# Get name for first
segment_names <- segment_names %>%
  rename(stop_sequence = first_stop_sequence)

segment_names <- left_join(segment_names, both_segments, by = c("route", "direction", "segment", "stop_sequence"))

segment_names <- segment_names %>%
  rename(first_stop_sequence = stop_sequence, first_stop_location = stop_location)

# Get name for last
segment_names <- segment_names %>%
  rename(stop_sequence = last_stop_sequence)

segment_names <- left_join(segment_names, both_segments, by = c("route", "direction", "segment", "stop_sequence"))

segment_names <- segment_names %>%
  rename(last_stop_sequence = stop_sequence, last_stop_location = stop_location) %>%
  select(-first_stop_sequence, -last_stop_sequence)

# Remove duplicates
segment_names <- segment_names %>%
  group_by(route, direction, segment) %>%
  summarise(first_stop_location = min(first_stop_location), last_stop_location = min(last_stop_location))

remove(champ_rail_segments, champ_name, both_segments)

```


#### Row-column method
```{r balancing-methods}
Fill_Failed_Output <- function(select_route, select_direction, select_time_of_day) {
  
  output_df <- data.frame(route = c(select_route), 
                          direction = c(select_direction),
                          time_of_day = c(select_time_of_day),
                          board_segment = c(NA),
                          alight_segment = c(NA),
                          flow = c(NA))
  
  return(output_df)
  
}

Balancing_Method <- function(select_route, select_direction, select_time_of_day, apc_df, champ_df) {
  
  # Build the row marginals (boardings)
  boardings_df <- apc_df %>%
    ungroup() %>%
    filter(route == select_route) %>%
    filter(direction == select_direction) %>%
    filter(time_of_day == select_time_of_day) %>%
    arrange(segment) %>%
    select(boardings)
  
  # Build the column marginals (alightings)
  alightings_df <- apc_df %>%
    ungroup() %>%
    filter(route == select_route) %>%
    filter(direction == select_direction) %>%
    filter(time_of_day == select_time_of_day) %>%
    arrange(segment) %>%
    select(alightings)
  
  # Build the seed matrix (champ flows)
  seed_df <- champ_df %>%
    ungroup() %>%
    filter(apc_route == select_route) %>%
    filter(apc_direction == select_direction) %>%
    filter(time_of_day == select_time_of_day) %>%
    select(board_segment, alight_segment, flow)
  
  mtx_rows = length(unique(seed_df$board_segment))
  mtx_cols = length(unique(seed_df$alight_segment))
  mtx_dims = max(mtx_rows, mtx_cols)
  
  # Fill in missing rows (TODO: make more elegant)
  segment_lookup = c("A", "B", "C", "D", "E", "F", "G")
  for (i in 1:mtx_dims) {
    
    row_segment = segment_lookup[i]
    
    for (j in 1:mtx_dims) {
      
      col_segment = segment_lookup[j]
      
      work_df <- seed_df %>%
        filter(board_segment  == row_segment) %>%
        filter(alight_segment == col_segment) 
      
      if (nrow(work_df) == 0) {
        board_segment  = c(row_segment)
        alight_segment = c(col_segment)
        
        # Use a seed of one unless it's going backwards
        flow = c(1L)
        if (i > j) flow = 0
        
        seed_df <- rbind(seed_df, data.frame(board_segment, alight_segment, flow))
        }
      
      } # j
    } # i
  
  seed_df <- seed_df %>%
    arrange(alight_segment, board_segment)
  
  seed_matrix <- matrix(seed_df$flow, mtx_dims, mtx_dims)
  
  # Fit the model
  fit_matrix <- loglin(outer(boardings_df$boardings, alightings_df$alightings) / sum(alightings_df$alightings), 
                       margin = list(1,2), 
                       start = seed_matrix, 
                       fit = TRUE, eps = THRESH, 
                       iter = ITER)$fit
  
  # Put the fit matrix back into a data frame
  output_df <- seed_df
  for (i in 1:mtx_dims) {
    boarding_segment = c(segment_lookup[i])
    for (j in 1:mtx_dims) {
      output_df <- output_df %>%
        mutate(flow = ifelse(board_segment == segment_lookup[i] & alight_segment == segment_lookup[j], fit_matrix[i,j], flow))
      }
    }
  
  output_df <- output_df %>%
    mutate(route = select_route) %>%
    mutate(direction = select_direction) %>%
    mutate(time_of_day = select_time_of_day) %>%
    filter(flow > 0)
  
  return(output_df)
  
  }

```

#### Apply method
```{r apply-method}
route_vector <- unique(apc_sum$route)
direction_vector <- unique(apc_sum$direction)
time_of_day_vector <- c("AM", "MD", "PM", "daily")

# First route
print(paste("Route: ", route_vector[1], 
            "  Direction: ", direction_vector[1], 
            "  Time of Day: ", time_of_day_vector[1], sep = ""))

output_df <- Balancing_Method(route_vector[1], 
                              direction_vector[1], 
                              time_of_day_vector[1],
                              apc_sum, champ_sum)

# Rest of the routes
for (i in 1:length(route_vector)) {
  route <- route_vector[i]
  
  for (j in 1:length(direction_vector)) {
    direction <- direction_vector[j]
    
    for (k in 1:length(time_of_day_vector)) {
      time_of_day <- time_of_day_vector[k]
      
      if (!(i == 1 & j == 1 & k == 1)) {
        
        print(paste("Route: ", route, 
                    "  Direction: ", direction, 
                    "  Time of Day: ", time_of_day, sep = ""))
        
        work_df <- tryCatch(Balancing_Method(route, direction,time_of_day,apc_sum, champ_sum),
                            error = function(e) Fill_Failed_Output(route, direction, time_of_day))
        
        output_df <- rbind(output_df, work_df)
        }
      
      } # k
    } # j
  } # i

# remove those without fits
output_df <- na.omit(output_df)

# add the segment names
output_df <- output_df %>%
  rename(segment = board_segment)
output_df <- left_join(output_df, segment_names, by = c("route", "direction", "segment"))
output_df <- output_df %>%
  rename(board_segment = segment, board_segment_start = first_stop_location, board_segment_end = last_stop_location)

output_df <- output_df %>%
  rename(segment = alight_segment)
output_df <- left_join(output_df, segment_names, by = c("route", "direction", "segment"))
output_df <- output_df %>%
  rename(alight_segment = segment, alight_segment_start = first_stop_location, alight_segment_end = last_stop_location)

remove(work_df)

```

#### Build comparisons
```{r comparison}
# Flow comparison that we will not write out
compare_champ <- champ_sum %>%
  rename(route = apc_route, direction = apc_direction, champ_flow = flow)

compare_champ <- left_join(output_df, 
                           compare_champ, 
                           by = c("route", "direction", "time_of_day", "board_segment", "alight_segment"))

# Missing champ_flows get a zero
compare_champ <- compare_champ %>%
  mutate(champ_flow = ifelse(is.na(champ_flow), 0, champ_flow))

# Summarise the output boardings and alightings by segment
one <- output_df %>%
  group_by(route, direction, time_of_day, board_segment) %>%
  summarise(boardings = sum(flow)) %>%
  rename(segment = board_segment)

two <- output_df %>%
  group_by(route, direction, time_of_day, alight_segment) %>%
  summarise(alightings = sum(flow)) %>%
  rename(segment = alight_segment)

three <- left_join(one, two, by = c("route", "direction", "time_of_day", "segment"))

# Join the output with the apc boardings and alightings by segment
compare_df <- apc_sum %>%
  rename(apc_boardings = boardings, apc_alightings = alightings)

compare_df <- left_join(three, compare_df, by = c("route", "direction", "time_of_day", "segment")) 

# Summarise the champ boardings and alightings by segment
one <- compare_champ %>%
  group_by(route, direction, time_of_day, board_segment) %>%
  summarise(champ_boardings = sum(champ_flow)) %>%
  rename(segment = board_segment)

two <- compare_champ %>%
  group_by(route, direction, time_of_day, alight_segment) %>%
  summarise(champ_alightings = sum(champ_flow)) %>%
  rename(segment = alight_segment)

three <- left_join(one, two, by = c("route", "direction", "time_of_day", "segment"))
compare_df <- left_join(compare_df, three, by = c("route", "direction", "time_of_day", "segment"))

compare_df[is.na(compare_df)] <- 0

# Add segment names
compare_df <- left_join(compare_df, segment_names, by = c("route", "direction", "segment"))
compare_df <- compare_df %>%
  rename(segment_start = first_stop_location, segment_end = last_stop_location)
  

remove(one, two, three)
```

#### Write output to disk
```{r write}
output_df <- output_df %>%
  select(route, direction, time_of_day, 
         board_segment, board_segment_start, board_segment_end, 
         alight_segment, alight_segment_start, alight_segment_end, 
         flow)

write.csv(output_df, file = F_OUTPUT_PRIORS, row.names = FALSE, quote = F)

compare_df <- compare_df %>%
  select(route, direction, time_of_day, segment, segment_start, segment_end, boardings, apc_boardings, champ_boardings, 
         alightings, apc_alightings, champ_alightings)

head(compare_df, n = 20)

write.csv(compare_df, file = F_OUTPUT_COMPS, row.names = FALSE, quote = F)
```
