---
title: "Muni Successive On-to-off"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes two rounds of on-to-off surveys on SF Muni. It compares the two data sets just as the `muni-travel-model-priors.Rmd` script compares the on-to-off counts with a set of priors derived from APC and travel model data.

#### Outputs
1.  A database of chi-squared test results and Hellinger distance calculations for use with Tableau workbook (`Compare Successive On-to-Off.twb`)

#### _TODO_
1. Refactor folder structure to make Muni specific

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(stringr)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Paramaters
```{r parameters}

# Build the data frame now, perhaps move to database if gets more complicated
obs_route_name = c("7-Haight/Noriega [ OUTBOUND ]",          "7-Haight/Noriega [ INBOUND ]",
                   "7R-Haight/Noriega Rapid [ OUTBOUND ]",   "7R-Haight/Noriega Rapid [ INBOUND ]",
                   "J-Church [ OUTBOUND ]",                  "J-Church [ INBOUND ]",
                   "KT-Ingleside/Third Street [ OUTBOUND ]", "KT-Ingleside/Third Street [ INBOUND ]",
                   "L-Taraval [ OUTBOUND ]",                 "L-Taraval [ INBOUND ]",
                   "M-Ocean View [ OUTBOUND ]",              "M-Ocean View [ INBOUND ]",
                   "N-Judah [ OUTBOUND ]",                   "N-Judah [ INBOUND ]")

route          = c("7",    "7",
                   "7R",   "7R",
                   "J",    "J",    
                   "KT",   "KT",   
                   "L",    "L",
                   "M",    "M",    
                   "N",    "N")


route_names <- data.frame(obs_route_name, route, stringsAsFactors = FALSE)

# Get direction
route_names <- route_names %>%
  mutate(inbound  = str_count(obs_route_name, "INBOUND")) %>%
  mutate(outbound = str_count(obs_route_name, "OUTBOUND")) %>%
  mutate(direction = "missing") %>%
  mutate(direction = ifelse(inbound == 1,  "INBOUND", direction)) %>%
  mutate(direction = ifelse(outbound == 1, "OUTBOUND", direction)) %>%
  select(-inbound, -outbound)

```

#### Remote I/O Locations
```{r remote-io}
F_OBS_FLOWS_ROUND_01 = "M:/Data/OnBoard/Data and Reports/Muni/On-to-Off/As CSV/MUNI_ON2OFF_DRAFT-FINAL_CONSOLIDATED_SUBMITTAL_1-7_20160727.csv"
F_OBS_FLOWS_ROUND_02 = "M:/Data/OnBoard/Data and Reports/Muni/On-to-Off/As CSV/MUNI_ON2OFF_FINAL_20161025_ROUND2.csv"

F_OUTPUT = "M:/Data/OnBoard/Data and Reports/Muni/On-to-Off/successive-on-to-off-results.csv"
```

#### Data reads
```{r data-reads}
obs_round_01 <- read.table(file = F_OBS_FLOWS_ROUND_01, header = TRUE, sep = ",", stringsAsFactors = FALSE)
obs_round_02 <- read.table(file = F_OBS_FLOWS_ROUND_02, header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

#### Data cleans
```{r data-clean}
working_01 <- obs_round_01 %>%
  mutate(round = 1L)

working_02 <- obs_round_02 %>%
  mutate(round = 2L)

working <- rbind(working_01, working_02)

working_sum <- working %>%
  select(round, route = ROUTE_DESCRIPTION, board_segment = SEGMENT_ON, alight_segment = SEGMENT_OFF, time_of_day = TIME.PERIOD) %>%
  group_by(round, route, time_of_day, board_segment, alight_segment) %>%
  summarise(records = n())

sum_round_02 <- working_sum %>%
  ungroup() %>%
  filter(round == 2) %>%
  select(-round) %>%
  rename(records_02 = records)

sum_round_01 <- working_sum %>%
  ungroup() %>%
  filter(round == 1) %>%
  select(-round) %>%
  rename(records_01 = records)

working_sum <- left_join(sum_round_02, sum_round_01, by = c("route", "time_of_day", "board_segment", "alight_segment"))

working_sum <- working_sum %>%
  mutate(records_01 = ifelse(is.na(records_01), 0L, records_01)) %>%
  mutate(records_02 = ifelse(is.na(records_02), 0L, records_02)) %>%
  rename(obs_route_name = route)

# Add route names and direction
working_sum <- left_join(working_sum, route_names, by = c("obs_route_name"))

working_sum <- working_sum %>%
  select(-obs_route_name)

# Rename round 01 prior and round 02 observed
working_sum <- working_sum %>%
  rename(observed_flow = records_02) %>%
  rename(prior_flow = records_01)

# Add daily time period
working_daily <- working_sum %>%
  group_by(route, direction, board_segment, alight_segment) %>%
  summarise(observed_flow = sum(observed_flow), prior_flow = sum(prior_flow)) %>%
  ungroup() %>%
  mutate(time_of_day = "DAILY")

working_sum <- rbind(working_sum, working_daily)

# compute the prior shares and prior estimate
temp_sum <- working_sum %>%
  group_by(route, direction, time_of_day) %>%
  summarise(prior_sum = sum(prior_flow), observed_sum = sum(observed_flow))

working_sum <- left_join(working_sum, temp_sum, by = c("route", "direction", "time_of_day"))

working_sum <- working_sum %>%
  mutate(prior_share = ifelse(prior_sum > 0.0, prior_flow / prior_sum, 0.0)) %>%
  mutate(prior_estimate = prior_share * observed_sum) %>%
  select(-prior_sum, -observed_sum, -prior_share)

# Fill in missing priors with 1
working_sum <- working_sum %>%
  mutate(prior_flow = ifelse(is.na(prior_flow), 1, prior_flow))

  
remove(working_01, working_02, working, sum_round_01, sum_round_02, temp_sum, working_daily)


```

#### Chi-squared tests
```{r chi-squared}
missing_priors <- working_sum %>%
  group_by(route, direction, time_of_day) %>%
  summarise(sum_prior_flow = sum(prior_flow)) %>%
  ungroup() %>%
  mutate(flag = ifelse(sum_prior_flow > 0, TRUE, FALSE)) %>%
  select(-sum_prior_flow)

ready_chi <- left_join(working_sum, missing_priors, by = c("route", "direction", "time_of_day"))

ready_chi <- ready_chi %>%
  filter(flag)

outcome_chi <- ready_chi %>%
  group_by(route, direction, time_of_day) %>%
  summarise(test_statistic  = chisq.test(observed_flow, p = prior_flow/sum(prior_flow))$statistic[["X-squared"]],
            test_df         = chisq.test(observed_flow, p = prior_flow/sum(prior_flow))$parameter[["df"]],
            test_p_value    = chisq.test(observed_flow, p = prior_flow/sum(prior_flow))$p.value) %>%
  ungroup()

remove(missing_priors)

```

#### Hellinger distance
```{r hellinger-dist}
# see discrete distributions formula here: https://en.wikipedia.org/wiki/Hellinger_distance

ready_hellinger <- ready_chi %>%
  select(route, direction, time_of_day, board_segment, alight_segment, observed_flow, prior_flow, prior_estimate)

flow_sums <- ready_hellinger %>%
  group_by(route, direction, time_of_day) %>%
  summarise(sum_observed_flow = sum(observed_flow), sum_prior_flow = sum(prior_flow))

hellinger_df <- left_join(ready_hellinger, flow_sums, by = c("route", "direction", "time_of_day"))
 
outcome_hellinger <- hellinger_df %>%
  mutate(obs_prob   = ifelse(sum_observed_flow > 0L, observed_flow / sum_observed_flow, 0L)) %>%
  mutate(prior_prob = ifelse(sum_prior_flow > 0L, prior_flow / sum_prior_flow, 0L)) %>%
  mutate(prob_diff = sqrt(obs_prob) - sqrt(prior_prob)) %>%
  mutate(prob_diff = prob_diff * prob_diff) %>%
  group_by(route, direction, time_of_day) %>%
  summarise(hellinger_dist = sum(prob_diff)) %>%
  mutate(hellinger_dist = 1/sqrt(2) * sqrt(hellinger_dist)) %>%
  ungroup()

remove(flow_sums, hellinger_df, ready_chi)

```


#### Prepare the outputs
```{r prep-outputs}
write <- left_join(ready_hellinger, outcome_chi, by = c("route", "direction", "time_of_day"))
write <- left_join(write, outcome_hellinger, by = c("route", "direction", "time_of_day"))

remove(outcome_chi, outcome_hellinger)
```

#### Data write
```{r data-write}
write.csv(write, file = F_OUTPUT, row.names = FALSE, quote = F)

```
