---
title: "Build Database"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes, modifies, and consolidates data extracted from SF MTA's standard APC summary worksheets.  An Excel macro found in `extract_flat_file.xlsm` is used to extract a route-specific database from the standard MTA route-specific spreadsheets.  These databases are consumed here.  

#### Outputs
1.  A consolidated database of APC data

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(reshape2)
suppressMessages(library(dplyr))
library(stringr)
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O Locations
```{r remote-io}
F_INPUT_DIRECTORY  = "M:/Data/Transit/2015 APC/Muni/"

F_OUTPUT = "M:/Data/Transit/2015 APC/Muni/consolidated-database.csv"

INPUT_FILE_PREFIX = "2015_Winter"

input_file_suffix_array = c("__NX","0001","01AX","01BX","0002",
                            "0003","0005","005L","0006","008X",
                            "08AX","08BX","0009","009L","0010",
                            "0012","0014","014L","014X","016X",
                            "0017","0018","0019","0021","0022",
                            "0023","0024","0027","0028","028L",
                            "0029","0030","030X","0031","31AX",
                            "31BX","0033","0035","0036","0037",
                            "0038","038L","38AX","0039","0041",
                            "0043","0044","0045","0047","0048",
                            "0049","0052","0054","0055","0056",
                            "0066","0067","0071","071L","081X",
                            "082X","083X","0088","0090","0091",
                            "0108")

```


#### Clean and re-shape data method
```{r clean-method}
clean_and_reshape_data <- function(input_file_suffix){
  
  # read in the data
  input_file_name <- paste(F_INPUT_DIRECTORY, INPUT_FILE_PREFIX, input_file_suffix, ".csv", sep = "")
  input_df <- read.table(file = input_file_name, header = TRUE, sep = ",", stringsAsFactors = FALSE, strip.white = TRUE)
  
  # cleaning step 1: variable tidying
  output_df <- input_df %>%
    filter(nchar(stop_location) > 3) %>%
    
    mutate(start_date = NA) %>%
    mutate(end_date   = NA) %>%
    mutate(start_date = ifelse(date_range == 'FEB02_MAR13_2015', as.Date("2015-02-02"), start_date)) %>%
    mutate(end_date   = ifelse(date_range == 'FEB02_MAR13_2015', as.Date("2015-03-13"), end_date))   %>%
    
    mutate(week_part = ifelse(week_part == "_WEEKDAYS", "WEEKDAYS", week_part)) %>%
    mutate(direction = ifelse(direction == "_INBOUND", "INBOUND", direction)) %>%
    mutate(route = paste(route)) %>%
    select(-date_range)
  
  # cleaning step 2: make variables from the strings
  output_df <- cbind(output_df, colsplit(output_df$stop_location, "-", c("stop_sequence", "stop_location_two")))
  
  output_df <- output_df %>%
    mutate(stop_sequence = as.numeric(stop_sequence)) %>%
    mutate(stop_location = stop_location_two) %>%
    select(-stop_location_two)
  
  output_df <- cbind(output_df, colsplit(output_df$capacity_str, "Capacity = ", c("delete_me", "capacity")))
  
  output_df <- output_df %>%
    mutate(capacity = as.numeric(capacity)) %>%
    select(-capacity_str, -delete_me)
  
  # reshape to get time dimension
  output_df <- melt(output_df, id = c("start_date", "end_date", "week_part", 
                                      "route", "direction", "stop_location", "stop_sequence", 
                                      "capacity_std", "capacity")) 
  
  output_df <- cbind(output_df, colsplit(output_df$variable, "_", c("on_off", "time_of_day")))
  
  output_df <- output_df %>%
    mutate(boardings = ifelse(on_off == "on", as.numeric(value), 0)) %>%
    mutate(alightings = ifelse(on_off == "off", as.numeric(value), 0)) %>%
    select(-variable, -value, -on_off)
  
  output_df <- output_df %>%
    group_by(start_date, end_date, week_part, route, direction, stop_location, stop_sequence, 
             time_of_day, capacity_std, capacity) %>%
    summarise(boardings = sum(boardings), alightings = sum(alightings))
  
}
```


#### Execute the method
```{r iterate-method}
for (i in 1:length(input_file_suffix_array)) {
  
  output_one <- clean_and_reshape_data(input_file_suffix_array[i])
  
  if (i == 1) output_bind <- output_one
  else output_bind <- rbind(output_bind, output_one)
  
}
```

#### Write to disk
```{r write}
write.csv(output_bind, file = F_OUTPUT, row.names = FALSE, quote = F)

```

