---
title: "Build-Production-Example-Inputs"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
good first start, need to iterate back once the production example is modified

#### Purpose
The next version of the production example (see production-example.Rmd) will read the following information from disk: 
(i) a survey file;
(ii) a csv defining the variables included in the observed targets;
(iii) a csv defining the observed targets count, importance, and keys.

Here, these files are prepared from the existing survey data. 

#### Overhead
```{r}
library(knitr)
library(stringr)
suppressMessages(library(dplyr))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data Preparation
#### Remote data locations
```{r data-reads}
input_dir  <- "M:/Data/OnBoard/Data and Reports/Tri Delta/As CSV/"
output_dir <- "D:/files/My Box Files/Share Data/multi-criteria-expansion/"

input.tablet <- read.table(file = paste(input_dir,"ONBOARD NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.onoff  <- read.table(file = paste(input_dir, "ON2OFF NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.sequence <- read.table(file = paste(input_dir, "RTD_ROUTE_STOPS_ALL_MODES_MAR_2 NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

#### Standardize data
```{r prepare-datasets}
# Survey data, align variables
survey <- input.tablet %>%
  select(tablet_id = ID, 
         route_dir = ROUTE.CODE, 
         direction = DIRECTION, 
         time_period = TIME_PERIOD,
         board_id = BOARDING_LOCATION_STOPID, 
         board_shp = BOARDING_LOCATION_STOPID_SHP, 
         alight_id = ALIGHTING_LOCATION_STOPID, 
         alight_shp = ALIGHTING_LOCATION_STOPID_SHP, 
         weight = UNLINKED_WGHT_FCTR) %>%
  mutate(route = as.numeric(str_sub(route_dir, 1, 3))) %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395) %>%
  mutate(time_period = ifelse(time_period == "LATE PM", "EVENING", time_period))

# Use on/off survey as a proxy for detailed observed data
observed <- input.onoff %>%
  select(onoff_id = ETC_ID, 
         route = ROUTE_ON, 
         direction = DIRECTION_ON, 
         time_period = TIME.PERIOD, 
         board_id = BOARDING_STOPID_, 
         board_shp = BOARDING_STOPID_SHP, 
         board_seq = BOARDING_STOPSEQUENCE, 
         alight_id = ALIGHTING_STOPID_, 
         alight_shp = ALIGHTING_STOPID_SHP, 
         alight_seq = ALIGHTING_STOPSEQUENCE, 
         match_confidence = matchConfidence) %>%
  mutate(time_period = ifelse(time_period == "AM1", "EARLY AM", time_period)) %>%
  mutate(time_period = ifelse(time_period == "AM2", "AM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "MID", "MIDDAY",   time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM1", "PM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM2", "EVENING",  time_period)) %>%
  filter(match_confidence == "100%")

stops <- input.sequence %>%
  select(operator = CPT_AGENCY, 
         route = SCH_ROUTED, 
         direction = RTD_PATTER, 
         stop_seq = SCH_STOPPO, 
         stop_shp = CPT_STOPPO) %>%
  filter(operator == '3D') %>%
  select(-operator)

````

#### Add route break points
```{r route-breaks}
# Determine maximum number of routes by direction
segment_breaks <- observed %>%
  group_by(route, direction) %>%
  summarise(max_board_seq = max(board_seq), max_alight_seq = max(alight_seq))

# TODO: make this more elegant/generalizable

# Create the segment breaks
segment_breaks <- segment_breaks %>%
  mutate(break_2_1 = round((max_board_seq + max_alight_seq) / 4)) %>%
  mutate(break_3_1 = round((max_board_seq + max_alight_seq) / 6)) %>%
  mutate(break_3_2 = break_3_1 * 2) %>%
  mutate(break_4_1 = round((max_board_seq + max_alight_seq) / 8)) %>%
  mutate(break_4_2 = break_4_1 * 2) %>%
  mutate(break_4_3 = break_4_1 * 3)

# Place the observed data into the segment breaks
observed <- left_join(observed, segment_breaks, by = c("route", "direction"))

observed <- observed %>%
  # Two segments
  mutate(board_seg_2  = ifelse(board_seq  < break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq < break_2_1, 1, 2)) %>%
  # Three segments
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  # Four segments
  mutate(board_seg_4  = ifelse(board_seq  <  break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq <  break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  # Remove excess variables
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

# Place the survey data into the segment breaks
stops_board <- stops %>%
  select(route, direction, board_seq = stop_seq, board_shp = stop_shp)

stops_alight <- stops %>%
  select(route, direction, alight_seq = stop_seq, alight_shp = stop_shp)

survey <- left_join(survey, stops_board, by = c("route",  "direction", "board_shp")) 
survey <- left_join(survey, stops_alight, by = c("route", "direction", "alight_shp"))
survey <- left_join(survey, segment_breaks, by = c("route", "direction"))

survey <- survey %>%
  select(-board_shp, -alight_shp)

survey <- survey %>%
  # Two segments
  mutate(board_seg_2  = ifelse(board_seq  < break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq < break_2_1, 1, 2)) %>%
  # Three segments
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  # Four segments
  mutate(board_seg_4  = ifelse(board_seq  <  break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq <  break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  # Remove excess variables
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

# TODO: what to do with non-matches (NA right now)
# should be fine as the survey record will not be incident to any targets if they are missing

```

#### Prepare the observed targets definition file
```{r observed-targets-defined}
observed_targets_defn <- data.frame(target_category_id = c("A", "B", "B", "B", "C", "C", "C", "D", "D", "D", "E", "E", "E", "E"), 
                                    survey_variable = c("all_routes",
                                                        "route", "direction", "time_period",
                                                        "route", "direction", "board_seg_4",
                                                        "route", "direction", "alight_seg_4",
                                                        "route", "direction", "board_seg_3", "alight_seg_3"))

```

#### Prepare observed target counts, importance weights, and target keys file
```{r observed-target-counts-weights}
# Set the importance weight variables
weight_A_all_routes = 20.0
weight_B_route_direction_time = 1.0
weight_C_route_direction_board  = 2.0
weight_D_route_direction_alight = 2.0
weight_E_route_direction_board_alight = 3.0

# Target Category A: all routes (from the survey)
observed_targets_A <- data.frame(target_count = sum(survey$weight))
observed_targets_A <- observed_targets_A %>%
  mutate(importance_weight = weight_A_all_routes) %>%
  mutate(target_category_id = "A") %>%
  mutate(route = NA) %>%
  mutate(direction = NA) %>%
  mutate(time_period  = NA) %>%
  mutate(board_seg_3  = NA) %>%
  mutate(alight_seg_3 = NA) %>%
  mutate(board_seg_4  = NA) %>%
  mutate(alight_seg_4 = NA)

# Target Category B: route, direction, time_period (from the survey using the existing weights)
observed_targets_B <- survey %>%
  group_by(route, direction, time_period) %>%
  summarise(target_count = sum(weight)) %>%
  ungroup() %>%
  mutate(importance_weight = weight_B_route_direction_time) %>%
  mutate(target_category_id = "B") %>%
  mutate(board_seg_3  = NA) %>%
  mutate(alight_seg_3 = NA) %>%
  mutate(board_seg_4  = NA) %>%
  mutate(alight_seg_4 = NA)

# Target Category C: route, direction, 4-category boarding segment  (from the on/off survey)
observed_targets_C <- observed %>%
  group_by(route, direction, board_seg_4) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(importance_weight = weight_C_route_direction_board) %>%
  mutate(target_category_id = "C") %>%
  mutate(time_period  = NA) %>%
  mutate(board_seg_3  = NA) %>%
  mutate(alight_seg_3 = NA) %>%
  mutate(alight_seg_4 = NA)

# Target Category D: route, direction, 4-category alighting segment  (from the on/off survey)
observed_targets_D <- observed %>%
  group_by(route, direction, alight_seg_4) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(importance_weight = weight_D_route_direction_alight) %>%
  mutate(target_category_id = "D") %>%
  mutate(time_period  = NA) %>%
  mutate(board_seg_3  = NA) %>%
  mutate(alight_seg_3 = NA) %>%
  mutate(board_seg_4  = NA)

# Target Category E: route, direction, 3-category boarding segment, 3-category alighting segment  (from the on/off survey)
observed_targets_E <- observed %>%
  group_by(route, direction, board_seg_3, alight_seg_3) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(importance_weight = weight_E_route_direction_board_alight) %>%
  mutate(target_category_id = "E") %>%
  mutate(time_period  = NA) %>%
  mutate(board_seg_4  = NA) %>%
  mutate(alight_seg_4 = NA)

# Select the variables needed for the counts and weights and bind with unique integer
observed_targets <- rbind(observed_targets_A, observed_targets_B, observed_targets_C, observed_targets_D, observed_targets_E)
target_id <- seq(1:length(observed_targets$target_category_id))
observed_targets <- cbind(target_id, observed_targets)
  
```

## Data writes
#### Write the example extracted input files to disk for others' reference
```{r data-writes}
write.csv(survey, file = paste(output_dir, "production-example-survey.csv",   sep = ""), row.names = FALSE, quote = F)
save(survey, file = paste(output_dir, "production-example-survey.Rdata", sep = ""))

write.csv(observed_targets_defn, file = paste(output_dir, "observed-target-definitions.csv", sep = ""), row.names = FALSE, quote = F)
save(observed_targets_defn, file = paste(output_dir, "observed-target-definitions.Rdata", sep = ""))

write.csv(observed_targets, file = paste(output_dir, "observed-target-counts.csv", sep = ""), row.names = FALSE, quote = F)
save(observed_targets, file = paste(output_dir, "observed-target-counts.Rdata", sep = ""))

```
