---
title: "Mini-Example"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
no known bugs

#### Purpose
Demonstrate multi-criteria transit on-board survey expansion using R tools via a mini (i.e., trivial) example. 

#### Overhead
```{r}
library(knitr)
suppressMessages(library(dplyr))
library(optimx)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Optimization Function
#### Inputs
1. `x` - a vector of length M representing each of the unique survey weights that need to be calculated
2. `obs_target_v` - a vector of length N containing each expansion target
3. `import_v` - a vector of length N containing the importance weight for each expansion target
4. `inc_mtx` - a matrix of dimensions M x N containing a dummy variable denoting the relevance of each unique survey weight to each expansion target
```{r optimization-function}
optimization_function <- function(x, obs_target_v, import_v, inc_mtx) {
  
  # Compute estimated targets 
  est_target_v <- x %*% inc_mtx
  
  # Compute importance-weighted errors, which is the objective function
  error_df <- data.frame(obs_target_v, est_target_v, import_v)
  error_df <- error_df %>%
    mutate(error = import_v * abs(est_target_v - obs_target_v))
  
  # Return errors
  return(sum(error_df$error))

}

```

## Mini-Example
Consider a survey with records for five passengers, A, B, C, D, and E.  Passengers A and B ride the 10E bus; passengers C, D, and E ride the 10W bus.  External information suggests the target ridership for the 10E bus is 100; the target ridership for the 10W bus is 300; and the system ridership is 450.
```{r mini-example}
#  Create the survey data
survey_id = c(1, 2, 3, 4, 5)
route = c(10, 10, 10, 10, 10)
direction = c("E", "E", "W", "W", "W")

survey_data = data.frame(survey_id, route, direction)

# Create the observed targets vector (length N, which is 3 in this example)
observed_targets_vector = c(target_1 = 100,
                            target_2 = 300,
                            target_3 = 450)

# Create the importance weights vector (assume we subjectively determine that matching target 3 is a bit more important than matching target 2 which is a bit more important than matching target 1)
importance_weights_vector = c(weight_1 = 1.00,
                              weight_2 = 1.25,
                              weight_3 = 1.50)

# Create the incidence matrix (length M, which is the number of unique weights (2), by N, the number of targets)
incidence_matrix = data.frame (target_1 = c(1, 0), # unique weight 1 (route 10E) is relevant to target 1
                               target_2 = c(0, 1), # unique weight 2 (route 10W) is relevant to target 2
                               target_3 = c(1, 1)) # both weights are relevant to target 3

incidence_matrix = data.matrix(incidence_matrix)

# Create a vector of starting weights
starting_weights_vector <- c(1,1)

# Run the optimization
optimx_results <- optimx(starting_weights_vector,
                         fn = optimization_function,
                         method = "L-BFGS-B",
                         lower = 1.00,
                         upper = Inf,
                         obs_target_v = observed_targets_vector,
                         import_v = importance_weights_vector,
                         inc_mtx = incidence_matrix)

# Extract and label results
unique_weights <- as.data.frame(t(coef(optimx_results)))
names(unique_weights)[1] <- "sum_weights"
unique_weights <- cbind(unique_weights, unique_weight_id = c("10---E", "10---W"))

# Summarize survey to unique categories and join optimal weights
survey_summary <- survey_data %>%
  group_by(route, direction) %>%
  summarise(records = n()) %>%
  mutate(unique_weight_id = paste(route, "---", direction, sep = ""))

survey_summary <- left_join(survey_summary, unique_weights, by = c("unique_weight_id"))

# Compute record weights and join to survey
survey_summary <- survey_summary %>%
  mutate(record_weight = sum_weights / records) %>%
  select(-unique_weight_id, -sum_weights, -records)

survey_data <- left_join(survey_data, survey_summary, by = c("route", "direction"))

survey_data


```


