---
title: "Production-Example"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
seems to be working; two examples in place

#### TODO
2. Vary weights, targets, etc, to make sure response is as expected (additional examples).

3. Catch inconsistent input errors

#### Purpose
Demonstrate multi-criteria transit on-board survey expansion using R tools via a production (i.e.,  practice-ready) example using real data.  Here, we use two surveys performed on Tri-Delta.  The first survey was a simple on-off count performed for ~70 percent of the riders.  The second is a personal interview survey of ~5 percent of the riders. Simple, external database files are used to define our targets.

## Key Input Files
#### Observed Target Counts
The `observed-target-counts` input database must have the following fields:
1. `target_id` - integer, a unique integer for the target;
2. `target_count` - float, the target value for the target (i.e., the weights will be adjusted to match the targets);
3. `importance_weight` - float, the importance weight for the target;
4. `target_category_id` - string, a label for each group of targets;
5. {survey variable names} - any number of columns that define the targets composition using the survey variable names, e.g., route, direction, time_period.  Each string MUST correspond to a variable in the survey data. 

#### Observed Target Definitions
The `observed-target-definitions` input database must have the following fields:
1. `target_category_id` - string, a label for each group of targets;
2. `survey_variable` - string, a label for the variables that comprise the target, with each variable being entered on a separate row in the database.  Each string MUST correspond to a variable in the survey data, with the exception of `all_routes` which denotes a target that is applied to all records in the survey.

Note that if a single, system-wide target is used, it must be in the first category listed.

#### Survey Data
The `survey` data must have the variables (as columns) defined in the `observed-target-counts` and `observed-target-definitions` files.

## Examples
#### Example 1: Five categories of targets
Please see `build-production-example-inputs.Rmd` for the creation of these files.  

Five categories of targets are defined, as follows:
1.  Category `A`: single, system-wide target;
2.  Category `B`: by route, direction, and time of day;
3.  Category `C`: by route, direction, and four-category boarding segment;
4.  Category `D`: by route, direction, and four-category alighting segment;
5.  Category `E`: by route, direction, three-category boarding segment, and three-category alighting segment.

Target categories `A` and `B` are drawn for the personal interview survey; categories `C`, `D`, and `E` are drawn from the on-off survey.

#### Example 2
This example is as subset of Example 1 in which we use these procedures to try and re-create the existing survey weights by expanding to route, direction, and time of day (i.e., Category `B` from Example 1).

#### Example 3
The same example as Example 1, but with different importance weights.  Specifically, we change:

1.  Category `A` importance weight from 20 to 20
2.  Category `B` importance weight from  1 to 5
3.  Category `C` importance weight from  2 to 0
4.  Category `D` importance weight from  2 to 0
5.  Category `E` importance weight from  3 to 1

We could, of course, remove categories `C` and `D`, but leave them in here to test the script mechanics.  We expect the RMSE performance of the different targets to move in the expected direction.

## Overhead

#### Libraries
```{r libraries}
source("method-library.R")
library(knitr)
suppressMessages(library(dplyr))
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data Preparation
#### Remote data locations
```{r data-reads}
remote_dir <- "D:/files/My Box Files/Share Data/multi-criteria-expansion/"
#remote_dir <- "~DavidWork/Documents/multi-criteria-expansion-inputs/"
```

## Convenience Methods
#### Root-mean-square-error
```{r rmse-method}
compute_target_rmse <- function(obs_targets_df, target_defn_df, survey_df, category_string){

  target <- obs_targets_df %>%
    filter(target_category_id == category_string)

  relevant_variables <- target_defn_df %>%
    filter(target_category_id == category_string)

  target_compare <- survey_df %>%
    group_by_(.dots = relevant_variables$survey_variable) %>%
    summarise(sum_base_weights = sum(weight), sum_optim_weights = sum(record_weight))

  target_compare <- left_join(target_compare, target, by = relevant_variables$survey_variable)

  target_compare <- target_compare %>%
    filter(!is.na(target_count)) %>%
    mutate(base_error  = sum_base_weights - target_count) %>%
    mutate(optim_error = sum_optim_weights - target_count) %>%
    mutate(base_error_sqd = base_error * base_error) %>%
    mutate(optim_error_sqd = optim_error * optim_error)

  return(c(sqrt(mean(target_compare$base_error_sqd)),
           sqrt(mean(target_compare$optim_error_sqd))))

}

```

#### Density plots
```{r density-plots}
make_a_density_plot <- function(survey_with_results_df, title_str){
  
  weight_data <- survey_with_results_df %>%
    select(weight) %>%
    mutate(type = "original")
  
  record_weight_data <- survey_with_results_df %>%
    select(weight = record_weight) %>%
    mutate(type = "optimal")
  
  weight_data <- rbind(weight_data, record_weight_data)
  
  ggplot(weight_data, aes (x = weight, fill = type, binwidth = 1.0)) + 
    geom_density(alpha = 0.4) + 
    labs(title = title_str) + 
    xlab("Survey Record Weight") + 
    ylab("Density")
  
  }

```


## Example 1
#### Example 1 Inputs
```{r example-01-inputs}
F_INPUT_SURVEY  <- paste(remote_dir, "production-example-survey.Rdata", sep = "")
F_INPUT_TARGETS <- paste(remote_dir, "observed-target-counts.Rdata", sep = "")
F_INPUT_DEFNS   <- paste(remote_dir, "observed-target-definitions.Rdata", sep = "")

load(F_INPUT_SURVEY)
load(F_INPUT_TARGETS)
load(F_INPUT_DEFNS)
```

#### Example 1 Paramaters
```{r example-01-parameters}
record_weight_lower_bound = 1.0
record_weight_upper_bound = 30.0
```

#### Example 1 Optimization
```{r example-01-optimization}
survey_with_results_ex01 <- execute_optimization(observed_targets, 
                                                 observed_targets_defn,
                                                 survey,
                                                 record_weight_lower_bound,
                                                 record_weight_upper_bound) 
```

#### Example 1 Summaries
```{r example-01-results}
# Print target RMSEs

# Target definitions variables must be characters
observed_targets_defn <- observed_targets_defn %>%
  mutate(target_category_id = as.character(target_category_id)) %>%
  mutate(survey_variable = as.character(survey_variable))

# target A
target <- observed_targets %>%
  filter(target_category_id == "A")
abs(sum(survey_with_results_ex01$weight) - sum(target$target_count))
abs(sum(survey_with_results_ex01$record_weight) - sum(target$target_count))

# target B
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex01, "B")
rmse_vector[1]
rmse_vector[2]

# target C
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex01, "C")
rmse_vector[1]
rmse_vector[2]

# target D
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex01, "D")
rmse_vector[1]
rmse_vector[2]

# target E
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex01, "E")
rmse_vector[1]
rmse_vector[2]

# Plot histrogram
example_01_density_plot <- make_a_density_plot(survey_with_results_ex01, 
                                               "Example 01 Distribution of Original and Optimal Weights")
example_01_density_plot

```

#### Example 1 clean-up
```{r example-01-clean-up}
survey_with_results_ex01 <- survey_with_results_ex01 %>%
  mutate(ex_01_weight = record_weight)
```


## Example 2
#### Example 2 Inputs
```{r example-02-inputs}
F_INPUT_SURVEY  <- paste(remote_dir, "production-example-survey.Rdata", sep = "")
F_INPUT_TARGETS <- paste(remote_dir, "observed-target-counts-ex02.csv", sep = "")
F_INPUT_DEFNS   <- paste(remote_dir, "observed-target-definitions-ex02.csv", sep = "")

load(F_INPUT_SURVEY)
observed_targets <- read.table(file = F_INPUT_TARGETS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
observed_targets_defn <- read.table(file = F_INPUT_DEFNS,   header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

#### Example 2 Paramaters
```{r example-02-parameters}
record_weight_lower_bound = 0.25
record_weight_upper_bound = 50.0 
```

#### Example 2 Optimization
```{r example-02-optimization}
survey_with_results_ex02 <- execute_optimization(observed_targets, 
                                                 observed_targets_defn,
                                                 survey,
                                                 record_weight_lower_bound,
                                                 record_weight_upper_bound) 
```


#### Example 2 Summaries
```{r example-02-results}
# Print target RMSEs

# target B
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex02, "B")
rmse_vector[1]
rmse_vector[2]

# Plot histrogram
example_02_density_plot <- make_a_density_plot(survey_with_results_ex02, 
                                               "Example 02 Distribution of Original and Optimal Weights")
example_02_density_plot

```

#### Example 2 clean-up
```{r example-02-clean-up}
survey_with_results_ex02 <- survey_with_results_ex02 %>%
  mutate(ex_02_weight = record_weight)
```


## Example 3
#### Example 3 Inputs
```{r example-03-inputs}
F_INPUT_SURVEY  <- paste(remote_dir, "production-example-survey.Rdata", sep = "")
F_INPUT_TARGETS <- paste(remote_dir, "observed-target-counts-ex03.csv", sep = "")
F_INPUT_DEFNS   <- paste(remote_dir, "observed-target-definitions-ex03.csv", sep = "")

load(F_INPUT_SURVEY)
observed_targets <- read.table(file = F_INPUT_TARGETS, header = TRUE, sep = ",", stringsAsFactors = FALSE)
observed_targets_defn <- read.table(file = F_INPUT_DEFNS,   header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

#### Example 3 Paramaters
```{r example-03-parameters}
record_weight_lower_bound = 1.0
record_weight_upper_bound = 30.0
```

#### Example 3 Optimization
```{r example-03-optimization}
survey_with_results_ex03 <- execute_optimization(observed_targets, 
                                                 observed_targets_defn,
                                                 survey,
                                                 record_weight_lower_bound,
                                                 record_weight_upper_bound) 
```

#### Example 3 Summaries
```{r example-03-results}
# Print target RMSEs

# Target definitions variables must be characters
observed_targets_defn <- observed_targets_defn %>%
  mutate(target_category_id = as.character(target_category_id)) %>%
  mutate(survey_variable = as.character(survey_variable))

# target A
target <- observed_targets %>%
  filter(target_category_id == "A")
abs(sum(survey_with_results_ex03$weight) - sum(target$target_count))
abs(sum(survey_with_results_ex03$record_weight) - sum(target$target_count))

# target B
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex03, "B")
rmse_vector[1]
rmse_vector[2]

# target C
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex03, "C")
rmse_vector[1]
rmse_vector[2]

# target D
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex03, "D")
rmse_vector[1]
rmse_vector[2]

# target E
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results_ex03, "E")
rmse_vector[1]
rmse_vector[2]

# Plot histrogram
example_03_density_plot <- make_a_density_plot(survey_with_results_ex03, 
                                               "Example 03 Distribution of Original and Optimal Weights")
example_03_density_plot

```

#### Example 3 clean-up
```{r example-03-clean-up}
survey_with_results_ex03 <- survey_with_results_ex03 %>%
  mutate(ex_03_weight = record_weight)
```


## Data Writes
#### Write the example outputs to disk
```{r data-writes}
# Example 1
write.csv(survey_with_results_ex01, 
          file = paste(remote_dir, "production-example-survey-results.csv",   sep = ""), 
          row.names = FALSE, 
          quote = F)

# Example 2
write.csv(survey_with_results_ex02, 
          file = paste(remote_dir, "production-example-02-survey-results.csv",   sep = ""), 
          row.names = FALSE, 
          quote = F)

# Example 3
write.csv(survey_with_results_ex03, 
          file = paste(remote_dir, "production-example-03-survey-results.csv",   sep = ""), 
          row.names = FALSE, 
          quote = F)

```

