---
title: "Production-Example"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
brute force approach *seems* to be working

#### TODO
1. Extract the route, direction, time of day targets, and the system-wide ridership target, from the survey's existing weights, to make the comparisons to the existing weights fair & informative.

2. Make the whole thing more elegant and generalizable (perhaps in a separate branch?), i.e. put the targets in a database in which parameters can be extracted for all the joining bits? make all the observed inputs external files (perhaps this is a branch?)? The current use case, in which we have all this observed data ready to go, is unlikely to be typical. If we do read in the key files externally, what checks are needed to make sure the optimization will work? Will the inputs point to variable names in the survey?  Think more about all of this.

3. Can we make the optimization faster?  Other methods?

4. Experiment with changing the target importance weights to make sure the response is as expected

#### Purpose
Demonstrate multi-criteria transit on-board survey expansion using R tools via a production (i.e.,  practice-ready) example using real data.  Here, we use two surveys performed on Tri-Delta.  The first survey was a simple on-off count performed for ~70 percent of the riders; we assume this data represents the population ridership characteristics.  The second survey is a personal interview survey containing typical questions; this survey has already been expanded to external data, which will give us some results to compare to.

#### Overhead
```{r}
library(knitr)
library(stringr)
library(optimx)
library(reshape2)
suppressMessages(library(dplyr))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data Preparation
#### Remote data locations
```{r data-reads}
input_dir  <- "M:/Data/OnBoard/Data and Reports/Tri Delta/As CSV/"
output_dir <- "D:/files/My Box Files/Share Data/multi-criteria-expansion/"

input.tablet <- read.table(file = paste(input_dir,"ONBOARD NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.onoff  <- read.table(file = paste(input_dir, "ON2OFF NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

input.sequence <- read.table(file = paste(input_dir, "RTD_ROUTE_STOPS_ALL_MODES_MAR_2 NO POUND OR SINGLE QUOTE.csv", sep = ""), header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

#### Standardize data
```{r prepare-datasets}
# Survey data, align variables
survey <- input.tablet %>%
  select(tablet_id = ID, 
         route_dir = ROUTE.CODE, 
         direction = DIRECTION, 
         time_period = TIME_PERIOD,
         board_id = BOARDING_LOCATION_STOPID, 
         board_shp = BOARDING_LOCATION_STOPID_SHP, 
         alight_id = ALIGHTING_LOCATION_STOPID, 
         alight_shp = ALIGHTING_LOCATION_STOPID_SHP, 
         weight = UNLINKED_WGHT_FCTR) %>%
  mutate(route = as.numeric(str_sub(route_dir, 1, 3))) %>%
  filter(route != 392) %>%
  filter(route != 393) %>%
  filter(route != 394) %>%
  filter(route != 395) %>%
  mutate(time_period = ifelse(time_period == "LATE PM", "EVENING", time_period))

# Use on/off survey as a proxy for detailed observed data
observed <- input.onoff %>%
  select(onoff_id = ETC_ID, 
         route = ROUTE_ON, 
         direction = DIRECTION_ON, 
         time_period = TIME.PERIOD, 
         board_id = BOARDING_STOPID_, 
         board_shp = BOARDING_STOPID_SHP, 
         board_seq = BOARDING_STOPSEQUENCE, 
         alight_id = ALIGHTING_STOPID_, 
         alight_shp = ALIGHTING_STOPID_SHP, 
         alight_seq = ALIGHTING_STOPSEQUENCE, 
         match_confidence = matchConfidence) %>%
  mutate(time_period = ifelse(time_period == "AM1", "EARLY AM", time_period)) %>%
  mutate(time_period = ifelse(time_period == "AM2", "AM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "MID", "MIDDAY",   time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM1", "PM PEAK",  time_period)) %>%
  mutate(time_period = ifelse(time_period == "PM2", "EVENING",  time_period)) %>%
  filter(match_confidence == "100%")

stops <- input.sequence %>%
  select(operator = CPT_AGENCY, 
         route = SCH_ROUTED, 
         direction = RTD_PATTER, 
         stop_seq = SCH_STOPPO, 
         stop_shp = CPT_STOPPO) %>%
  filter(operator == '3D') %>%
  select(-operator)

````

#### Add route break points
```{r route-breaks}
# Determine maximum number of routes by direction
segment_breaks <- observed %>%
  group_by(route, direction) %>%
  summarise(max_board_seq = max(board_seq), max_alight_seq = max(alight_seq))

# TODO: make this more elegant/generalizable

# Create the segment breaks
segment_breaks <- segment_breaks %>%
  mutate(break_2_1 = round((max_board_seq + max_alight_seq) / 4)) %>%
  mutate(break_3_1 = round((max_board_seq + max_alight_seq) / 6)) %>%
  mutate(break_3_2 = break_3_1 * 2) %>%
  mutate(break_4_1 = round((max_board_seq + max_alight_seq) / 8)) %>%
  mutate(break_4_2 = break_4_1 * 2) %>%
  mutate(break_4_3 = break_4_1 * 3)

# Place the observed data into the segment breaks
observed <- left_join(observed, segment_breaks, by = c("route", "direction"))

observed <- observed %>%
  # Two segments
  mutate(board_seg_2  = ifelse(board_seq  < break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq < break_2_1, 1, 2)) %>%
  # Three segments
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  # Four segments
  mutate(board_seg_4  = ifelse(board_seq  <  break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq <  break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  # Remove excess variables
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

# Place the survey data into the segment breaks
stops_board <- stops %>%
  select(route, direction, board_seq = stop_seq, board_shp = stop_shp)

stops_alight <- stops %>%
  select(route, direction, alight_seq = stop_seq, alight_shp = stop_shp)

survey <- left_join(survey, stops_board, by = c("route",  "direction", "board_shp")) 
survey <- left_join(survey, stops_alight, by = c("route", "direction", "alight_shp"))
survey <- left_join(survey, segment_breaks, by = c("route", "direction"))

survey <- survey %>%
  select(-board_shp, -alight_shp)

survey <- survey %>%
  # Two segments
  mutate(board_seg_2  = ifelse(board_seq  < break_2_1, 1, 2)) %>%
  mutate(alight_seg_2 = ifelse(alight_seq < break_2_1, 1, 2)) %>%
  # Three segments
  mutate(board_seg_3  = ifelse(board_seq  <  break_3_1, 1, 2)) %>%
  mutate(board_seg_3  = ifelse(board_seq  >= break_3_2, 3, board_seg_3)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq <  break_3_1, 1, 2)) %>%
  mutate(alight_seg_3 = ifelse(alight_seq >= break_3_2, 3, alight_seg_3)) %>%
  # Four segments
  mutate(board_seg_4  = ifelse(board_seq  <  break_4_1, 1, 2)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_1 & board_seq < break_4_2, 2, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_2 & board_seq < break_4_3, 3, board_seg_4)) %>%
  mutate(board_seg_4  = ifelse(board_seq  >= break_4_3, 4, board_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq <  break_4_1, 1, 2)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_1 & alight_seq < break_4_2, 2, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_2 & alight_seq < break_4_3, 3, alight_seg_4)) %>%
  mutate(alight_seg_4 = ifelse(alight_seq >= break_4_3, 4, alight_seg_4)) %>%
  # Remove excess variables
  select(-break_2_1, -break_3_1, -break_3_2, -break_4_1, -break_4_2, -break_4_3, -max_board_seq, -max_alight_seq)

# TODO: what to do with non-matches (NA right now)
# should be fine as the survey record will not be incident to any targets if they are missing

```


## Optimization Function
#### Inputs
1. `x` - a vector of length M representing each of the unique survey weights that need to be calculated
2. `obs_target_v` - a vector of length N containing each expansion target
3. `import_v` - a vector of length N containing the importance weight for each expansion target
4. `inc_mtx` - a matrix of dimensions M x N containing a dummy variable denoting the relevance of each unique survey weight to each expansion target
```{r optimization-function}
optimization_function <- function(x, obs_target_v, import_v, inc_mtx) {
  
  # Compute estimated targets 
  est_target_v <- x %*% inc_mtx
  
  # Compute importance-weighted errors, which is the objective function
  error_df <- data.frame(obs_target_v, est_target_v, import_v)
  error_df <- error_df %>%
    mutate(error <- import_v * abs(est_target_v - obs_target_v))
  
  # Return errors
  return(sum(error_df$error))

}

```

## Production-Example
Expand the Tri-Delta survey by (i) route, direction, and time-of-day; (ii) system-wide ridership; (iii) route, direction, and boarding sequence category (4 categories); (iv) route, direction, and alighting sequence category (4 categories); and, (v) route, direction, boarding sequence category, and alighting sequence category (3 categories).  

#### Prepare the observed targets vector
```{r observed-targets}
# Target Set 1: Route, direction, and time-of-day
observed_targets_1 <- observed %>%
  group_by(route, direction, time_period) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(target_id = paste(route, "---", direction, "---", time_period, sep = "")) %>%
  select(target_id, target_count)

# Target Set 2: System-wide ridership (match sum of existing survey weights)
observed_targets_2 <- data.frame(target_id = c("all_routes"), target_count = c(9849))

# Target Set 3: Route, direction, 4-category boarding segment
observed_targets_3 <- observed %>%
  group_by(route, direction, board_seg_4) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(target_id = paste(route, "---", direction, "---board-", board_seg_4, sep = "")) %>%
  select(target_id, target_count)

# Target Set 4: Route, direction, 4-category alighting segment
observed_targets_4 <- observed %>%
  group_by(route, direction, alight_seg_4) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(target_id = paste(route, "---", direction, "---alight-", alight_seg_4, sep = "")) %>%
  select(target_id, target_count)

# Target Set 5: Route, direction, 3-category boarding, 3-category alighting
observed_targets_5 <- observed %>%
  group_by(route, direction, board_seg_3, alight_seg_3) %>%
  summarise(target_count = n()) %>%
  ungroup() %>%
  mutate(target_id = paste(route, "---", direction, "---board-", board_seg_3, "---alight-", alight_seg_3, sep = "")) %>%
  select(target_id, target_count)

# Combine the targets
observed_targets <- rbind(observed_targets_1, observed_targets_2, observed_targets_3, observed_targets_4, observed_targets_5)
observed_targets_vector <- observed_targets$target_count

```

#### Build the importance weight vector
```{r importance-weights}
weight_route_direction_time = 1.0
weight_all_routes = 20.0
weight_route_direction_board  = 2.0
weight_route_direction_alight = 2.0
weight_route_direction_board_alight = 3.0

importance_weights <- observed_targets %>%
  # Default weight
  mutate(weight = weight_route_direction_time) %>%
  # All routes
  mutate(weight = ifelse(str_detect(target_id, "all_routes"), weight_all_routes, weight)) %>%
  # Boardings
  mutate(weight = ifelse(str_detect(target_id, "board")  & !str_detect(target_id, "alight"), weight_route_direction_board, weight)) %>%
  # Alightings
  mutate(weight = ifelse(!str_detect(target_id, "board") & str_detect(target_id, "alight"), weight_route_direction_alight, weight)) %>%
  # Boardings + Alightings
  mutate(weight = ifelse(str_detect(target_id, "board")  & str_detect(target_id, "alight"), weight_route_direction_board_alight, weight))

importance_weights_vector <- importance_weights$weight

```

#### Build the incidence matrix
```{r incidence-matrix}
# TODO: make more elegant/generalizable

# Create a data set with the unique weights we'll need for the survey
unique_weights <- survey %>%
  group_by(route, direction, time_period, board_seg_3, board_seg_4, alight_seg_3, alight_seg_4) %>%
  summarise(records = n()) %>%
  mutate(target_id_1 = paste(route, "---", direction, "---", time_period, sep = "")) %>%
  mutate(target_id_2 = "all_routes") %>%
  mutate(target_id_3 = paste(route, "---", direction, "---board-", board_seg_4, sep = "")) %>%
  mutate(target_id_4 = paste(route, "---", direction, "---alight-", alight_seg_4, sep = "")) %>%
  mutate(target_id_5 = paste(route, "---", direction, "---board-", board_seg_3, "---alight-", alight_seg_3, sep = ""))

# Append the observed targets incidence matrices, one at a time

# Target 1 (route, direction, time period)
observed_targets_incidence_1 <- observed_targets_1 %>%
  mutate(one = 1)
observed_targets_incidence_1 <- dcast(observed_targets_incidence_1, target_id ~ target_id, value.var = "one")
observed_targets_incidence_1[is.na(observed_targets_incidence_1)] <- 0
observed_targets_incidence_1 <- observed_targets_incidence_1 %>%
  rename(target_id_1 = target_id)

# Target 2 (all routes)
observed_targets_incidence_2 <- observed_targets_2 %>%
  mutate(one = 1)
observed_targets_incidence_2 <- dcast(observed_targets_incidence_2, target_id ~ target_id, value.var = "one")
observed_targets_incidence_2[is.na(observed_targets_incidence_2)] <- 0
observed_targets_incidence_2 <- observed_targets_incidence_2 %>%
  rename(target_id_2 = target_id)

# Target 3 (route, direction, boarding segment)
observed_targets_incidence_3 <- observed_targets_3 %>%
  mutate(one = 1)
observed_targets_incidence_3 <- dcast(observed_targets_incidence_3, target_id ~ target_id, value.var = "one")
observed_targets_incidence_3[is.na(observed_targets_incidence_3)] <- 0
observed_targets_incidence_3 <- observed_targets_incidence_3 %>%
  rename(target_id_3 = target_id)

# Target 4 (route, direction, alighting segment)
observed_targets_incidence_4 <- observed_targets_4 %>%
  mutate(one = 1)
observed_targets_incidence_4 <- dcast(observed_targets_incidence_4, target_id ~ target_id, value.var = "one")
observed_targets_incidence_4[is.na(observed_targets_incidence_4)] <- 0
observed_targets_incidence_4 <- observed_targets_incidence_4 %>%
  rename(target_id_4 = target_id)

# Target 5 (route, direction, boarding segment, alighting segment)
observed_targets_incidence_5 <- observed_targets_5 %>%
  mutate(one = 1)
observed_targets_incidence_5 <- dcast(observed_targets_incidence_5, target_id ~ target_id, value.var = "one")
observed_targets_incidence_5[is.na(observed_targets_incidence_5)] <- 0
observed_targets_incidence_5 <- observed_targets_incidence_5 %>%
  rename(target_id_5 = target_id)


# Join the incidence matrices to the survey
incidence_matrix <- unique_weights %>%
  ungroup() %>%
  select(target_id_1, target_id_2, target_id_3, target_id_4, target_id_5)

incidence_matrix <- left_join(incidence_matrix, observed_targets_incidence_1, by = c("target_id_1"))
incidence_matrix <- left_join(incidence_matrix, observed_targets_incidence_2, by = c("target_id_2"))
incidence_matrix <- left_join(incidence_matrix, observed_targets_incidence_3, by = c("target_id_3"))
incidence_matrix <- left_join(incidence_matrix, observed_targets_incidence_4, by = c("target_id_4"))
incidence_matrix <- left_join(incidence_matrix, observed_targets_incidence_5, by = c("target_id_5"))

incidence_matrix <- incidence_matrix %>%
  select(-target_id_1) %>%
  select(-target_id_2) %>%
  select(-target_id_3) %>%
  select(-target_id_4) %>%
  select(-target_id_5)

# Fill the NAs with 0
incidence_matrix[is.na(incidence_matrix)] <- 0

incidence_matrix <- data.matrix(incidence_matrix)
  
```

#### Execute the optimization and extract results
```{r run-optimization}
# Create a vector of starting weights
starting_weights_vector <- unique_weights$records * 0 + 1.0

# Test the optimization function
optimization_function(starting_weights_vector, observed_targets_vector, importance_weights_vector, incidence_matrix)

# Run the optimization (keep track of computing time)
start_time <- proc.time()
optimx_results <- optimx(starting_weights_vector,
                         fn = optimization_function,
                         method = "L-BFGS-B",
                         lower = 1.00,
                         upper = Inf,
                         obs_target_v = observed_targets_vector,
                         import_v = importance_weights_vector,
                         inc_mtx = incidence_matrix)
proc.time() - start_time
# ~2200 seconds (~35 minutes)

sum_unique_weights <- as.data.frame(t(coef(optimx_results)))
names(sum_unique_weights)[1] <- "sum_weights"

```

#### Join optimization results to survey data
```{r join-results}
survey_summary <- cbind(unique_weights, sum_unique_weights)

survey_summary <- survey_summary %>%
  mutate(record_weight = sum_weights / records) %>%
  select(-target_id_1, -target_id_2, -target_id_3, -target_id_4, -target_id_5, -sum_weights, -records)

survey <- left_join(survey, survey_summary, by = c("route", "direction", "time_period", "board_seg_3", "board_seg_4", "alight_seg_3", "alight_seg_4"))

```

#### Explore the difference in the two sets of weights
```{r weight-diff}
survey_explore <- survey %>%
  mutate(diff = record_weight - weight) %>%
  mutate(abs_diff = abs(diff)) %>%
  arrange(desc(abs_diff))

# Compare the two sets of weights against each of the targets
# Note this is not an apples to apples comparison as the base weights are expanded to external data, not the on/off survey

# Target 1: route, direction, time
target_compare <- survey %>%
  group_by(route, direction, time_period) %>%
  summarise(sum_base_weights = sum(weight), sum_optim_weights = sum(record_weight)) %>%
  mutate(target_id = paste(route, "---", direction, "---", time_period, sep = ""))

target_compare <- left_join(target_compare, observed_targets_1, by = c("target_id"))

target_compare <- target_compare %>%
  mutate(base_error  = sum_base_weights - target_count) %>%
  mutate(optim_error = sum_optim_weights - target_count) %>%
  mutate(base_error_sqd = base_error * base_error) %>%
  mutate(optim_error_sqd = optim_error * optim_error) %>%
  arrange(desc(base_error_sqd))

sqrt(mean(target_compare$base_error_sqd))
sqrt(mean(target_compare$optim_error_sqd))

# Base: 121; Optim: 54

# Target 2: system wide ridership
abs(sum(survey$weight)-sum(observed_targets_2$target_count))
abs(sum(survey$record_weight)-sum(observed_targets_2$target_count))

# Base: 1842; Optim: 0

# TODO: check why the base is so far off (external counts likely inconsistent with )

# Target 3: route, direction, boarding location
target_compare <- survey %>%
  group_by(route, direction, board_seg_4) %>%
  summarise(sum_base_weights = sum(weight), sum_optim_weights = sum(record_weight)) %>%
  mutate(target_id = paste(route, "---", direction, "---board-", board_seg_4, sep = ""))

target_compare <- left_join(target_compare, observed_targets_3, by = c("target_id"))

target_compare <- target_compare %>%
  filter(!is.na(target_count))

target_compare <- target_compare %>%
  mutate(base_error  = sum_base_weights - target_count) %>%
  mutate(optim_error = sum_optim_weights - target_count) %>%
  mutate(base_error_sqd = base_error * base_error) %>%
  mutate(optim_error_sqd = optim_error * optim_error) %>%
  arrange(desc(base_error_sqd))

sqrt(mean(target_compare$base_error_sqd))
sqrt(mean(target_compare$optim_error_sqd))

# Base: 154; Optim: 25

# Target 4: route, direction, alighting location
target_compare <- survey %>%
  group_by(route, direction, alight_seg_4) %>%
  summarise(sum_base_weights = sum(weight), sum_optim_weights = sum(record_weight)) %>%
  mutate(target_id = paste(route, "---", direction, "---alight-", alight_seg_4, sep = ""))

target_compare <- left_join(target_compare, observed_targets_4, by = c("target_id"))

target_compare <- target_compare %>%
  filter(!is.na(target_count))

target_compare <- target_compare %>%
  mutate(base_error  = sum_base_weights - target_count) %>%
  mutate(optim_error = sum_optim_weights - target_count) %>%
  mutate(base_error_sqd = base_error * base_error) %>%
  mutate(optim_error_sqd = optim_error * optim_error) %>%
  arrange(desc(base_error_sqd))

sqrt(mean(target_compare$base_error_sqd))
sqrt(mean(target_compare$optim_error_sqd))
  
# Base: 139; Optim: 33

# Target 5: route, direction, boarding, alighting
target_compare <- survey %>%
  group_by(route, direction, board_seg_3, alight_seg_3) %>%
  summarise(sum_base_weights = sum(weight), sum_optim_weights = sum(record_weight)) %>%
  mutate(target_id = paste(route, "---", direction, "---board-", board_seg_3, "---alight-", alight_seg_3, sep = ""))

target_compare <- left_join(target_compare, observed_targets_5, by = c("target_id"))

target_compare <- target_compare %>%
  filter(!is.na(target_count))

target_compare <- target_compare %>%
  mutate(base_error  = sum_base_weights - target_count) %>%
  mutate(optim_error = sum_optim_weights - target_count) %>%
  mutate(base_error_sqd = base_error * base_error) %>%
  mutate(optim_error_sqd = optim_error * optim_error) %>%
  arrange(desc(base_error_sqd))

sqrt(mean(target_compare$base_error_sqd))
sqrt(mean(target_compare$optim_error_sqd))

# Base: 98; Optim: 15
```

## Data writes
#### Write the example extracted input files to disk for others' reference
```{r data-writes}
write.csv(survey,   file = paste(output_dir, "production-example-survey.csv",   sep = ""), row.names = FALSE, quote = F)
write.csv(observed, file = paste(output_dir, "production-example-observed.csv", sep = ""), row.names = FALSE, quote = F)

```

