---
title: "Caltrain-Expansion"
author: "David Ory"
html_document:
    toc: true
    theme: cosmo
runtime: shiny
---

## Administration

#### Status
seems to be working, though the optimization currently takes 20 hours

#### TODO
0. is run time a function of wildly inconsistent targets? re-scale target C to find out

1. when done, merge into standard database?

2. narrow the range of the bounds so it runs faster

3. rebuild the targets with the on/off counts from the sampling plan?


#### Purpose
The Caltrain survey was expanded by the consultant two ways: (a) by boarding station, alighting station, and time of day; and, (b) by route, time, and direction.  These two expansions were done independently.

Here, we use the multi-criteria transit expansion procedures to expand the survey using the following targets:

1.  Category `A`: single, system-wide target;
2.  Category `B`: by boarding station, alighting station, and time-of-day;
3.  Category `C`: by train number (each of which is unique).
4.  Category `D`: by service type (local, limited, baby bullet)

See `build-caltrain-expansion-inputs.Rmd` for the cross-walk from the survey to the target inputs.

## Overhead

#### Libraries
```{r libraries}
source("../method-library.R")
library(knitr)
suppressMessages(library(dplyr))
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data Preparation
#### Remote data locations
```{r data-reads}
input_dir <- "M:/Data/OnBoard/Data and Reports/Caltrain/" 

F_SURVEY  <- paste(input_dir, "As CSV/Caltrain_Final_Submitted_1_5_2015_withSERVICETYPE NO POUND OR SINGLE QUOTE.csv", sep = "")
F_TARGETS <- paste(input_dir, "multi-criteria-expansion/observed-target-counts.Rdata", sep = "")
F_DEFNS   <- paste(input_dir, "multi-criteria-expansion/observed-target-definitions.Rdata", sep = "")

survey <- read.table(file = F_SURVEY, header = TRUE, sep = ",", stringsAsFactors = FALSE)
load(F_TARGETS)
load(F_DEFNS)

```

#### Standardize the survey variable names
```{r survey-standard}
survey <- survey %>%
  select(survey_id = ID, 
         board_station = ENTER_STATION,
         alight_station = EXIT_STATION,
         time_period = TIME_PERIOD,
         line_number = RUN,
         service_type = SERVICE_TYPE,
         weight_bat = LINKED_WGT_FCTR_BtoA,
         weight_rtd = LINKED_WGT_FCTR_RTD) %>%
  filter(!is.na(weight_bat))
         
```

## Optimization
```{r optimization}
record_weight_lower_bound = 1.0
record_weight_upper_bound = 40.0

survey_with_results <- execute_optimization(observed_targets, 
                                            observed_targets_defn,
                                            survey,
                                            record_weight_lower_bound,
                                            record_weight_upper_bound) 
```

```{r debug}
# target_counts_df <- observed_targets
# targets_defn_df <- observed_targets_defn
# survey_df <- survey
# rw_lower_scalar <- record_weight_lower_bound
# rw_upper_scalar <- record_weight_upper_bound
# 

```

## Convenience methods
#### Root-mean-square-error
```{r rmse-method}
compute_target_rmse <- function(obs_targets_df, target_defn_df, survey_df, category_string){

  target <- obs_targets_df %>%
    filter(target_category_id == category_string)

  relevant_variables <- target_defn_df %>%
    filter(target_category_id == category_string)

  target_compare <- survey_df %>%
    group_by_(.dots = relevant_variables$survey_variable) %>%
    summarise(sum_bat_weights = sum(weight_bat), 
              sum_rtd_weights = sum(weight_rtd),
              sum_opt_weights = sum(record_weight))

  target_compare <- left_join(target_compare, target, by = relevant_variables$survey_variable)

  target_compare <- target_compare %>%
    filter(!is.na(target_count)) %>%
    mutate(bat_error = sum_bat_weights - target_count) %>%
    mutate(rtd_error = sum_rtd_weights - target_count) %>%
    mutate(opt_error = sum_opt_weights - target_count) %>%
    mutate(bat_error_sqd = bat_error * bat_error) %>%
    mutate(rtd_error_sqd = rtd_error * rtd_error) %>%
    mutate(opt_error_sqd = opt_error * opt_error)

  return(c(sqrt(mean(target_compare$bat_error_sqd)),
           sqrt(mean(target_compare$rtd_error_sqd)),
           sqrt(mean(target_compare$opt_error_sqd))))

}
```

## Summarize results
```{r assessment}
# Target definitions variables must be characters
observed_targets_defn <- observed_targets_defn %>%
  mutate(target_category_id = as.character(target_category_id)) %>%
  mutate(survey_variable = as.character(survey_variable))

# Target A
target <- observed_targets %>%
  filter(target_category_id == "A")
cat("Original B/A: ", abs(sum(survey_with_results$weight_bat) - sum(target$target_count)))
cat("Original RTD: ", abs(sum(survey_with_results$weight_rtd) - sum(target$target_count)))
cat("Optimal     : ", abs(sum(survey_with_results$record_weight) - sum(target$target_count)))

# Target A
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results, "A")
rmse_vector[1]
rmse_vector[2]
rmse_vector[3]

# Target B
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results, "B")
rmse_vector[1]
rmse_vector[2]
rmse_vector[3]

# Target C
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results, "C")
rmse_vector[1]
rmse_vector[2]
rmse_vector[3]

# Target D
rmse_vector <- compute_target_rmse(observed_targets, observed_targets_defn, survey_with_results, "D")
rmse_vector[1]
rmse_vector[2]
rmse_vector[3]

# Plot the distribution of the three weight sets
bat_weight <- survey_with_results %>%
  select(weight = weight_bat) %>%
  mutate(type = "original b/a")

rtd_weight <- survey_with_results %>%
  select(weight = weight_rtd) %>%
  mutate(type = "original rtd")

opt_weight <- survey_with_results %>%
  select(weight = record_weight) %>%
  mutate(type = "optimal")

chart_data <- rbind(bat_weight, rtd_weight, opt_weight)

ggplot(chart_data, aes (x = weight, fill = type, binwidth = 1.0)) + 
  geom_density(alpha = 0.4) + 
  labs(title = "Caltrain Expansion -- Three Ways") + 
  xlab("Survey Record Weight") + 
  ylab("Density")

```

